---
title: "04 - Bayesian Logistic Regression"
output: html_document
---
# Overview

This script implements the Bayesian logistic regression analysis used in the thesis.

The script:
- loads the imputed datasets,
- fits full Bayesian logistic regression models using HMC,
- performs spike-and-slab variable selection,
- refits the final Bayesian models,
- computes posterior summaries, diagnostics, predicted probabilities, and predictive performance,
- evaluates computational runtime,
- generates figures.

```{r}
# ==================================================
# Load packages, paths, and imputed datasets
# ==================================================

library(mice)
library(brms)
library(dplyr)
library(purrr)
library(tibble)
library(tidyr)
library(posterior)
library(nimble)
library(ggplot2)
library(pROC)
library(knitr)

options(brms.backend = "rstan")

output_path <- "."

bayes_results_path <- file.path(
  output_path,
  "Results",
  "Bayesian Approach"
)

fig_path <- file.path(output_path, "figures")
tables_path <- file.path(output_path, "Tables")

dir.create(bayes_results_path, recursive = TRUE, showWarnings = FALSE)
dir.create(fig_path, recursive = TRUE, showWarnings = FALSE)
dir.create(tables_path, recursive = TRUE, showWarnings = FALSE)

imp_m5 <- readRDS(
  file.path(output_path, "Imputed-Data", "imputed_data_m5.rds")
)

imp_m10 <- readRDS(
  file.path(output_path, "Imputed-Data", "imputed_data_m10.rds")
)
```


```{r}
# ==================================================
# Extract completed datasets
# ==================================================

completed_m5  <- complete(imp_m5, action = "all")
completed_m10 <- complete(imp_m10, action = "all")

```


```{r}
# ==================================================
# Define Bayesian model, priors, and MCMC settings
# ==================================================

full_formula <- bf(
  HeartDisease ~ Age + Sex + ChestPainType + RestingBP + Cholesterol +
    FastingBS + RestingECG + MaxHR + ExerciseAngina + Oldpeak + ST_Slope
)

priors <- c(
  prior(student_t(7, 0, 2.5), class = "b"),
  prior(student_t(7, 0, 10), class = "Intercept")
)

bayes_chains <- 4
bayes_cores <- min(4, parallel::detectCores())
bayes_iter <- 800
bayes_warmup <- 400
bayes_seed <- 123

# ==================================================
# Fit full Bayesian logistic regression models
# ==================================================

fit_full_bayes <- function(completed_data_list) {
  
  map(
    completed_data_list,
    ~ brm(
      formula = full_formula,
      data = .x,
      family = bernoulli(link = "logit"),
      prior = priors,
      chains = bayes_chains,
      iter = bayes_iter,
      warmup = bayes_warmup,
      cores = bayes_cores,
      seed = bayes_seed,
      control = list(adapt_delta = 0.95)
    )
  )
}

fits_full_m5 <- fit_full_bayes(completed_m5)

fits_full_m10 <- fit_full_bayes(completed_m10)

```



```{r}
# ==================================================
# Pool posterior draws across imputations
# ==================================================

pool_bayes_draws <- function(fits) {
  
  draws <- map2_dfr(
    fits,
    seq_along(fits),
    function(fit, imp) {
      as_draws_df(fit) %>%
        as.data.frame() %>%
        select(any_of("b_Intercept"), starts_with("b_")) %>%
        mutate(imputation = imp)
    }
  )
  
  draws %>%
    select(-imputation) %>%
    pivot_longer(
      cols = everything(),
      names_to = "term",
      values_to = "estimate"
    ) %>%
    group_by(term) %>%
    summarise(
      posterior_mean = mean(estimate),
      lower_95 = quantile(estimate, 0.025),
      upper_95 = quantile(estimate, 0.975),
      OR = exp(posterior_mean),
      OR_lower = exp(lower_95),
      OR_upper = exp(upper_95),
      .groups = "drop"
    )
}

pooled_full_m5 <- pool_bayes_draws(fits_full_m5)
pooled_full_m10 <- pool_bayes_draws(fits_full_m10)

# Save pooled posterior summaries

write.csv(
  pooled_full_m5,
  file.path(bayes_results_path, "bayesian_full_model_m5.csv"),
  row.names = FALSE
)

write.csv(
  pooled_full_m10,
  file.path(bayes_results_path, "bayesian_full_model_m10.csv"),
  row.names = FALSE
)
```



```{r}
# ==================================================
# Extract HMC diagnostics
# ==================================================

extract_hmc_diagnostics <- function(fits) {
  
  map2_dfr(
    fits,
    seq_along(fits),
    function(fit, imp) {
      
      posterior::summarise_draws(
        as_draws_array(fit),
        rhat,
        ess_bulk,
        ess_tail,
        mcse_mean,
        sd
      ) %>%
        as.data.frame() %>%
        filter(grepl("^b_|^Intercept", variable)) %>%
        mutate(
          imputation = imp,
          mcse_ratio = mcse_mean / sd
        )
    }
  )
}

hmc_diagnostics_m5 <- extract_hmc_diagnostics(fits_full_m5)
hmc_diagnostics_m10 <- extract_hmc_diagnostics(fits_full_m10)

```


```{r}
# ==================================================
# Save HMC diagnostics
# ==================================================

write.csv(
  hmc_diagnostics_m5,
  file.path(bayes_results_path, "bayesian_hmc_diagnostics_m5.csv"),
  row.names = FALSE
)

write.csv(
  hmc_diagnostics_m10,
 file.path(bayes_results_path, "bayesian_hmc_diagnostics_m10.csv"),
  row.names = FALSE
)
```



```{r}
# =======================================================
# Run spike-and-slab variable selection and compute PIPs
# =======================================================

full_formula_selection <- HeartDisease ~ Age + Sex + ChestPainType + RestingBP + Cholesterol +
  FastingBS + RestingECG + MaxHR + ExerciseAngina + Oldpeak + ST_Slope

run_spike_slab_selection <- function(dat, formula_full, niter = 8000, nburnin = 2000) {
  
  X <- model.matrix(formula_full, data = dat)[, -1]
  y <- as.numeric(as.character(dat$HeartDisease))
  
  N <- nrow(X)
  P <- ncol(X)
  
  code <- nimbleCode({
    for (i in 1:N) {
      logit(p[i]) <- beta0 + inprod(beta_included[1:P], X[i, 1:P])
      y[i] ~ dbern(p[i])
    }
    
    beta0 ~ dnorm(0, sd = 5)
    
    for (j in 1:P) {
      gamma[j] ~ dbern(0.5)
      beta[j] ~ dnorm(0, sd = 2.5)
      beta_included[j] <- gamma[j] * beta[j]
    }
  })
  
  model <- nimbleModel(
    code = code,
    constants = list(N = N, P = P, X = X),
    data = list(y = y),
    inits = list(beta0 = 0, beta = rep(0, P), gamma = rep(1, P))
  )
  
  conf <- configureMCMC(model, monitors = c("gamma"))
  mcmc <- buildMCMC(conf)
  
  cmodel <- compileNimble(model)
  cmcmc <- compileNimble(mcmc, project = model)
  
  samples <- runMCMC(
    cmcmc,
    niter = niter,
    nburnin = nburnin,
    samplesAsCodaMCMC = FALSE
  )
  
  gamma_samples <- samples[, grep("^gamma", colnames(samples))]
  
  tibble(
    term = colnames(X),
    pip = colMeans(gamma_samples)
  )
}

# Run spike-and-slab variable selection 

spike_slab_pips_m5 <- map(
  completed_m5,
  run_spike_slab_selection,
  formula_full = full_formula_selection
)

spike_slab_pips_m10 <- map(
  completed_m10,
  run_spike_slab_selection,
  formula_full = full_formula_selection
)

```


```{r}
# =============================================================
# Average posterior inclusion probabilities across imputations
# =============================================================

avg_pips_m5 <- bind_rows(spike_slab_pips_m5, .id = "imputation") %>%
  group_by(term) %>%
  summarise(
    mean_pip = mean(pip),
    sd_pip = sd(pip),
    .groups = "drop"
  ) %>%
  arrange(desc(mean_pip))

avg_pips_m10 <- bind_rows(spike_slab_pips_m10, .id = "imputation") %>%
  group_by(term) %>%
  summarise(
    mean_pip = mean(pip),
    sd_pip = sd(pip),
    .groups = "drop"
  ) %>%
  arrange(desc(mean_pip))

```


```{r}
# ==================================================
# Save posterior inclusion probabilities
# ==================================================

write.csv(
  avg_pips_m5,
  file.path(bayes_results_path, "bayesian_pips_m5.csv"),
  row.names = FALSE
)

write.csv(
  avg_pips_m10,
  file.path(bayes_results_path, "bayesian_pips_m10.csv"),
  row.names = FALSE
)
```



```{r}
# ==================================================
# Select variables using PIP threshold
# ==================================================

clean_variable_names <- function(terms) {
  terms %>%
    gsub("^Sex.*", "Sex", .) %>%
    gsub("^ChestPainType.*", "ChestPainType", .) %>%
    gsub("^FastingBS.*", "FastingBS", .) %>%
    gsub("^RestingECG.*", "RestingECG", .) %>%
    gsub("^ExerciseAngina.*", "ExerciseAngina", .) %>%
    gsub("^ST_Slope.*", "ST_Slope", .) %>%
    unique()
}

selected_bayes_m5_raw <- avg_pips_m5 %>%
  filter(mean_pip > 0.5) %>%
  pull(term)

selected_bayes_m10_raw <- avg_pips_m10 %>%
  filter(mean_pip > 0.5) %>%
  pull(term)

selected_bayes_m5 <- clean_variable_names(selected_bayes_m5_raw)
selected_bayes_m10 <- clean_variable_names(selected_bayes_m10_raw)

```


```{r}
# ==================================================
# Refit final Bayesian models 
# ==================================================

final_formula_m5 <- as.formula(
  paste("HeartDisease ~", paste(selected_bayes_m5, collapse = " + "))
)

final_formula_m10 <- as.formula(
  paste("HeartDisease ~", paste(selected_bayes_m10, collapse = " + "))
)

# Use the same selected predictors for both final Bayesian models

original_variable_order <- c(
  "Age", "Sex", "ChestPainType", "RestingBP", "Cholesterol",
  "FastingBS", "RestingECG", "MaxHR", "ExerciseAngina",
  "Oldpeak", "ST_Slope"
)

selected_bayes_common <- original_variable_order[
  original_variable_order %in% selected_bayes_m5 &
    original_variable_order %in% selected_bayes_m10
]


final_formula_bayes <- as.formula(
  paste("HeartDisease ~", paste(selected_bayes_common, collapse = " + "))
)


fit_final_bayes <- function(completed_data_list, formula_final) {
  
  map(
    completed_data_list,
    ~ brm(
      formula = formula_final,
      data = .x,
      family = bernoulli(link = "logit"),
      prior = priors,
      chains = bayes_chains,
      iter = bayes_iter,
      warmup = bayes_warmup,
      cores = bayes_cores,
      seed = bayes_seed,
      control = list(adapt_delta = 0.95)
    )
  )
}


fits_final_m5 <- fit_final_bayes(completed_m5, final_formula_bayes)

fits_final_m10 <- fit_final_bayes(completed_m10, final_formula_bayes)

```


```{r}
# ==================================================
# Extract HMC diagnostics for final Bayesian models
# ==================================================

hmc_diagnostics_final_m5 <- extract_hmc_diagnostics(fits_final_m5)
hmc_diagnostics_final_m10 <- extract_hmc_diagnostics(fits_final_m10)

# Save HMC diagnostics for final Bayesian models

write.csv(
  hmc_diagnostics_final_m5,
  file.path(bayes_results_path, "bayesian_hmc_diagnostics_final_m5.csv"),
  row.names = FALSE
)

write.csv(
  hmc_diagnostics_final_m10,
  file.path(bayes_results_path, "bayesian_hmc_diagnostics_final_m10.csv"),
  row.names = FALSE
)
```



```{r}
# =================================================================
# Compute Bayesian predicted probabilities for each imputed dataset
# =================================================================

get_bayesian_predicted_probabilities <- function(fits_final, m_value) {
  
  map_dfr(
    fits_final,
    function(fit) {
      data.frame(
        observation = seq_len(nrow(fit$data)),
        HeartDisease = fit$data$HeartDisease,
        predicted_probability = fitted(fit, scale = "response")[, "Estimate"]
      )
    },
    .id = "imputation"
  ) %>%
    mutate(
      imputation = as.numeric(imputation),
      m = m_value
    )
}


bayesian_predicted_probabilities_m5 <- get_bayesian_predicted_probabilities(
  fits_final = fits_final_m5,
  m_value = 5
)

bayesian_predicted_probabilities_m10 <- get_bayesian_predicted_probabilities(
  fits_final = fits_final_m10,
  m_value = 10
)


# Averaged predicted probabilities for plots 

bayesian_predicted_probabilities_avg_m5 <- bayesian_predicted_probabilities_m5 %>%
  group_by(observation, HeartDisease) %>%
  summarise(
    predicted_probability = mean(predicted_probability),
    .groups = "drop"
  ) %>%
  mutate(m = 5)

bayesian_predicted_probabilities_avg_m10 <- bayesian_predicted_probabilities_m10 %>%
  group_by(observation, HeartDisease) %>%
  summarise(
    predicted_probability = mean(predicted_probability),
    .groups = "drop"
  ) %>%
  mutate(m = 10)

write.csv(
  bayesian_predicted_probabilities_m5,
  file.path(bayes_results_path, "bayesian_predicted_probabilities_by_imputation_m5.csv"),
  row.names = FALSE
)

write.csv(
  bayesian_predicted_probabilities_m10,
  file.path(bayes_results_path, "bayesian_predicted_probabilities_by_imputation_m10.csv"),
  row.names = FALSE
)

write.csv(
  bayesian_predicted_probabilities_avg_m5,
  file.path(bayes_results_path, "bayesian_predicted_probabilities_avg_m5.csv"),
  row.names = FALSE
)

write.csv(
  bayesian_predicted_probabilities_avg_m10,
  file.path(bayes_results_path, "bayesian_predicted_probabilities_avg_m10.csv"),
  row.names = FALSE
)
```



```{r}
# =============================================================
# Bayesian predictive performance with 95% confidence intervals
# =============================================================

compute_bayesian_performance_by_imputation <- function(pred_data, m_value) {
  
  pred_data <- pred_data %>%
    mutate(
      HeartDisease = as.numeric(as.character(HeartDisease)),
      predicted_class = ifelse(predicted_probability >= 0.5, 1, 0)
    )
  
  pred_data %>%
    group_by(imputation) %>%
    summarise(
      AUC = as.numeric(
        auc(
          roc(
            response = HeartDisease,
            predictor = predicted_probability,
            quiet = TRUE
          )
        )
      ),
      
      Accuracy = mean(predicted_class == HeartDisease),
      
      TP = sum(predicted_class == 1 & HeartDisease == 1),
      TN = sum(predicted_class == 0 & HeartDisease == 0),
      FP = sum(predicted_class == 1 & HeartDisease == 0),
      FN = sum(predicted_class == 0 & HeartDisease == 1),
      
      Sensitivity = ifelse((TP + FN) == 0, NA, TP / (TP + FN)),
      Specificity = ifelse((TN + FP) == 0, NA, TN / (TN + FP)),
      
      .groups = "drop"
    ) %>%
    mutate(m = m_value)
}


mean_ci <- function(x) {
  
  x <- x[!is.na(x)]
  
  m <- mean(x)
  se <- sd(x) / sqrt(length(x))
  
  c(
    mean = m,
    lower = m - 1.96 * se,
    upper = m + 1.96 * se
  )
}


summarise_bayesian_performance <- function(performance_by_imp, m_value) {
  
  performance_by_imp %>%
    summarise(
      m = m_value,
      n_imputations = n(),
      
      AUC_mean = mean_ci(AUC)["mean"],
      AUC_lower = mean_ci(AUC)["lower"],
      AUC_upper = mean_ci(AUC)["upper"],
      
      Accuracy_mean = mean_ci(Accuracy)["mean"],
      Accuracy_lower = mean_ci(Accuracy)["lower"],
      Accuracy_upper = mean_ci(Accuracy)["upper"],
      
      Sensitivity_mean = mean_ci(Sensitivity)["mean"],
      Sensitivity_lower = mean_ci(Sensitivity)["lower"],
      Sensitivity_upper = mean_ci(Sensitivity)["upper"],
      
      Specificity_mean = mean_ci(Specificity)["mean"],
      Specificity_lower = mean_ci(Specificity)["lower"],
      Specificity_upper = mean_ci(Specificity)["upper"]
    )
}
```



```{r}
# =======================================================
# Compute performance separately for each imputed dataset
# =======================================================

bayes_performance_by_imp_m5 <- compute_bayesian_performance_by_imputation(
  pred_data = bayesian_predicted_probabilities_m5,
  m_value = 5
)

bayes_performance_by_imp_m10 <- compute_bayesian_performance_by_imputation(
  pred_data = bayesian_predicted_probabilities_m10,
  m_value = 10
)


# Summarise mean performance with 95% confidence intervals

bayes_performance_summary_m5 <- summarise_bayesian_performance(
  bayes_performance_by_imp_m5,
  m_value = 5
)

bayes_performance_summary_m10 <- summarise_bayesian_performance(
  bayes_performance_by_imp_m10,
  m_value = 10
)


bayes_performance_results <- bind_rows(
  bayes_performance_summary_m5,
  bayes_performance_summary_m10
)

bayes_performance_results
```


```{r}
# ==================================================
# Plot ROC curve for Bayesian model m = 10
# ==================================================

bayes_roc_m10 <- roc(
  response = bayesian_predicted_probabilities_avg_m10$HeartDisease,
  predictor = bayesian_predicted_probabilities_avg_m10$predicted_probability,
  quiet = TRUE
)

roc_plot_bayes_m10 <- ggroc(
  bayes_roc_m10,
  legacy.axes = TRUE,
  linewidth = 1.4
) +
  geom_abline(
    intercept = 0,
    slope = 1,
    linetype = "dashed",
    colour = "grey60"
  ) +
  labs(
    title = "Bayesian ROC Curve",
    subtitle = paste0(
      "Final Bayesian model for m = 10 (AUC = ",
      round(bayes_performance_summary_m10$AUC_mean, 3),
      ")"
    ),
    x = "1 - Specificity",
    y = "Sensitivity"
  ) +
  coord_cartesian(
    xlim = c(0, 1),
    ylim = c(0, 1)
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(face = "bold", size = 18),
    plot.subtitle = element_text(size = 14),
    axis.title = element_text(face = "bold", size = 15),
    axis.text = element_text(size = 13),
    panel.grid.minor = element_blank()
  )

roc_plot_bayes_m10

ggsave(
  filename = file.path(fig_path, "bayesian_roc_curve_m10.png"),
  plot = roc_plot_bayes_m10,
  width = 8,
  height = 5.5,
  dpi = 300
)
```


```{r}
# ==================================================
# Save Bayesian predictive performance results
# ==================================================

write.csv(
  bayesian_predicted_probabilities_m5,
  file.path(bayes_results_path, "bayesian_predicted_probabilities_by_imputation_m5.csv"),
  row.names = FALSE
)

write.csv(
  bayesian_predicted_probabilities_m10,
  file.path(bayes_results_path, "bayesian_predicted_probabilities_by_imputation_m10.csv"),
  row.names = FALSE
)

write.csv(
  bayesian_predicted_probabilities_avg_m5,
  file.path(bayes_results_path, "bayesian_predicted_probabilities_avg_m5.csv"),
  row.names = FALSE
)

write.csv(
  bayesian_predicted_probabilities_avg_m10,
  file.path(bayes_results_path, "bayesian_predicted_probabilities_avg_m10.csv"),
  row.names = FALSE
)

write.csv(
  bayes_performance_by_imp_m5,
  file.path(bayes_results_path, "bayesian_predictive_performance_by_imputation_m5.csv"),
  row.names = FALSE
)

write.csv(
  bayes_performance_by_imp_m10,
  file.path(bayes_results_path, "bayesian_predictive_performance_by_imputation_m10.csv"),
  row.names = FALSE
)

write.csv(
  bayes_performance_results,
  file.path(bayes_results_path, "bayesian_predictive_performance_summary_CI.csv"),
  row.names = FALSE
)

```


```{r}
# ==================================================
# Pool final Bayesian posterior draws
# ==================================================

pooled_final_m5 <- pool_bayes_draws(fits_final_m5)
pooled_final_m10 <- pool_bayes_draws(fits_final_m10)

```


```{r}
# ==================================================
# Save final Bayesian model results
# ==================================================

write.csv(
  pooled_final_m5,
  file.path(bayes_results_path, "bayesian_final_results_m5.csv"),
  row.names = FALSE
)

write.csv(
  pooled_final_m10,
  file.path(bayes_results_path, "bayesian_final_results_m10.csv"),
  row.names = FALSE
)
```


```{r}
# ==================================================
# Create Bayesian summary tables
# ==================================================

all_original_vars <- c(
  "Age", "Sex", "ChestPainType", "RestingBP", "Cholesterol",
  "FastingBS", "RestingECG", "MaxHR", "ExerciseAngina",
  "Oldpeak", "ST_Slope"
)

make_bayesian_results <- function(
    pooled_full,
    pooled_final,
    selected_bayes,
    fits_final,
    m_value
) {
  
  selected_vars_clean <- clean_variable_names(selected_bayes)
  removed_bayes <- setdiff(all_original_vars, selected_vars_clean)
  
  bayesian_full_results <- pooled_full %>%
    mutate(
      term = gsub("^b_", "", term),
      interval_width = OR_upper - OR_lower
    )
  
  bayesian_final_results <- pooled_final %>%
    mutate(
      term = gsub("^b_", "", term),
      interval_width = OR_upper - OR_lower
    )
  
  bayesian_model_summary <- data.frame(
    m = m_value,
    selected_variables = paste(selected_vars_clean, collapse = ", "),
    number_selected_variables = length(selected_vars_clean),
    removed_variables = paste(removed_bayes, collapse = ", "),
    number_removed_variables = length(removed_bayes)
  )
  
  bayesian_posterior_probabilities <- map_dfr(
    fits_final,
    function(fit) {
      as_draws_df(fit) %>%
        as.data.frame() %>%
        select(starts_with("b_")) %>%
        summarise(across(everything(), ~ mean(.x > 0))) %>%
        pivot_longer(
          cols = everything(),
          names_to = "term",
          values_to = "P_beta_greater_0"
        )
    },
    .id = "imputation"
  ) %>%
    group_by(term) %>%
    summarise(
      mean_P_beta_greater_0 = mean(P_beta_greater_0),
      .groups = "drop"
    ) %>%
    mutate(term = gsub("^b_", "", term))
  
  bayesian_comparison_final_results <- bayesian_final_results %>%
    left_join(
      bayesian_posterior_probabilities,
      by = "term"
    ) %>%
    mutate(
      m = m_value,
      selected_variables = paste(selected_vars_clean, collapse = ", "),
      number_selected_variables = length(selected_vars_clean),
      removed_variables = paste(removed_bayes, collapse = ", "),
      number_removed_variables = length(removed_bayes)
    )
  
  list(
    full_results = bayesian_full_results,
    final_results = bayesian_final_results,
    posterior_probabilities = bayesian_posterior_probabilities,
    model_summary = bayesian_model_summary,
    comparison_final_results = bayesian_comparison_final_results
  )
}

bayesian_results_m5 <- make_bayesian_results(
  pooled_full = pooled_full_m5,
  pooled_final = pooled_final_m5,
  selected_bayes = selected_bayes_m5,
  fits_final = fits_final_m5,
  m_value = 5
)

bayesian_results_m10 <- make_bayesian_results(
  pooled_full = pooled_full_m10,
  pooled_final = pooled_final_m10,
  selected_bayes = selected_bayes_m10,
  fits_final = fits_final_m10,
  m_value = 10
)

bayesian_model_summary_m5_m10 <- bind_rows(
  bayesian_results_m5$model_summary,
  bayesian_results_m10$model_summary
)

bayesian_comparison_final_results_m5_m10 <- bind_rows(
  bayesian_results_m5$comparison_final_results,
  bayesian_results_m10$comparison_final_results
)

```



```{r}
# ==================================================
# Save Bayesian summary tables
# ==================================================

write.csv(
  bayesian_model_summary_m5_m10,
  file.path(bayes_results_path, "bayesian_model_summary_m5_m10.csv"),
  row.names = FALSE
)

write.csv(
  bayesian_comparison_final_results_m5_m10,
  file.path(bayes_results_path, "bayesian_comparison_final_results_m5_m10.csv"),
  row.names = FALSE
)
```



```{r}
# ==================================================
# Bayesian runtime stability
# ==================================================

n_time_runs <- 5

time_bayes_full_model <- function(run_id, completed_data_list, m_value) {
  
  time_out <- system.time({
    fit_full_bayes(completed_data_list)
  })
  
  tibble(
    procedure = "Full Bayesian HMC model",
    m = m_value,
    run = run_id,
    elapsed_time_seconds = as.numeric(time_out["elapsed"])
  )
}

time_bayes_selection <- function(run_id, completed_data_list, m_value) {
  
  time_out <- system.time({
    map(
      completed_data_list,
      run_spike_slab_selection,
      formula_full = full_formula_selection
    )
  })
  
  tibble(
    procedure = "Bayesian spike-and-slab variable selection",
    m = m_value,
    run = run_id,
    elapsed_time_seconds = as.numeric(time_out["elapsed"])
  )
}

time_bayes_final_model <- function(run_id, completed_data_list, formula_final, m_value) {
  
  time_out <- system.time({
    fit_final_bayes(completed_data_list, formula_final)
  })
  
  tibble(
    procedure = "Final Bayesian HMC model",
    m = m_value,
    run = run_id,
    elapsed_time_seconds = as.numeric(time_out["elapsed"])
  )
}

time_bayes_prediction <- function(run_id, fits_final, m_value) {
  
  time_out <- system.time({
    map_dfr(
      fits_final,
      function(fit) {
        data.frame(
          observation = seq_len(nrow(fit$data)),
          HeartDisease = fit$data$HeartDisease,
          predicted_probability = fitted(fit)[, "Estimate"]
        )
      },
      .id = "imputation"
    ) %>%
      group_by(observation, HeartDisease) %>%
      summarise(
        predicted_probability = mean(predicted_probability),
        .groups = "drop"
      )
  })
  
  tibble(
    procedure = "Bayesian predicted probabilities",
    m = m_value,
    run = run_id,
    elapsed_time_seconds = as.numeric(time_out["elapsed"])
  )
}

bayes_runtime_stability_raw <- bind_rows(
  
  map_dfr(1:n_time_runs, ~ time_bayes_full_model(.x, completed_m5, 5)),
  
  map_dfr(1:n_time_runs, ~ time_bayes_selection(.x, completed_m5, 5)),
  
  map_dfr(1:n_time_runs, ~ time_bayes_final_model(.x, completed_m5, final_formula_m5, 5)),
  
  map_dfr(1:n_time_runs, ~ time_bayes_prediction(.x, fits_final_m5, 5))
)

bayes_runtime_stability_summary <- bayes_runtime_stability_raw %>%
  group_by(procedure, m) %>%
  summarise(
    Mean_Runtime = paste0(
      round(mean(elapsed_time_seconds, na.rm = TRUE), 3),
      " ± ",
      round(sd(elapsed_time_seconds, na.rm = TRUE), 3),
      " sec"
    ),
    .groups = "drop"
  )

write.csv(
  bayes_runtime_stability_raw,
  file.path(bayes_results_path, "bayesian_runtime_stability_raw.csv"),
  row.names = FALSE
)

write.csv(
  bayes_runtime_stability_summary,
  file.path(bayes_results_path, "bayesian_runtime_stability_summary.csv"),
  row.names = FALSE
)
```




```{r}
# ==================================================
# Create Bayesian Plots
# ==================================================

# Create Bayesian PIP plot

to_original_variable <- function(x) {
  x %>%
    gsub("^Sex.*", "Sex", .) %>%
    gsub("^ChestPainType.*", "ChestPainType", .) %>%
    gsub("^FastingBS.*", "FastingBS", .) %>%
    gsub("^RestingECG.*", "RestingECG", .) %>%
    gsub("^ExerciseAngina.*", "ExerciseAngina", .) %>%
    gsub("^ST_Slope.*", "ST_Slope", .)
}

pips_original_m5 <- avg_pips_m5 %>%
  mutate(variable = to_original_variable(term)) %>%
  group_by(variable) %>%
  summarise(mean_pip = max(mean_pip), .groups = "drop")

pips_original_m10 <- avg_pips_m10 %>%
  mutate(variable = to_original_variable(term)) %>%
  group_by(variable) %>%
  summarise(mean_pip = max(mean_pip), .groups = "drop")

pip_original_plot <- bind_rows(
  pips_original_m5 %>% mutate(m = "m = 5"),
  pips_original_m10 %>% mutate(m = "m = 10")
) %>%
  mutate(variable = reorder(variable, mean_pip)) %>%
  ggplot(aes(x = variable, y = mean_pip, fill = m)) +
  geom_col(position = "dodge") +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  coord_flip() +
  labs(
    title = "Bayesian Variable Selection",
    subtitle = "Posterior inclusion probabilities",
    x = "Variable",
    y = "Posterior inclusion probability",
    fill = "Imputations"
  ) +
  theme_minimal(base_size = 16) +
theme(
  plot.title = element_text(
    size = 18,
    face = "bold",
    hjust = 0.5
  ),
  plot.subtitle = element_text(
    size = 15,
    hjust = 0.5
  ),
  axis.title = element_text(
    size = 16,
    face = "bold"
  ),
  axis.text = element_text(
    size = 16
  ),
  legend.title = element_text(
    size = 15,
    face = "bold"
  ),
  legend.text = element_text(
    size = 14
  ),
  panel.grid.minor = element_blank()
)

pip_original_plot

ggsave(
  file.path(fig_path, "bayesian_pip_original_variables.png"),
  pip_original_plot,
  width = 14,
  height = 8,
  dpi = 300
)


# Create Bayesian forest plot

clean_term <- function(x) {
  x %>%
    gsub("^b_", "", .) %>%
    gsub("^ChestPainType", "ChestPainType: ", .) %>%
    gsub("^ST_Slope", "ST_Slope: ", .) %>%
    gsub("^SexM$", "Sex: Male", .) %>%
    gsub("^FastingBS1$", "FastingBS: Yes", .) %>%
    gsub("^ExerciseAnginaY$", "ExerciseAngina: Yes", .)
}

bayesian_forest_plot <- bayesian_results_m10$final_results %>%
  filter(term != "Intercept") %>%
  mutate(
    term = clean_term(term),
    term = reorder(term, OR)
  ) %>%
  ggplot(aes(x = OR, y = term)) +
  geom_point() +
  geom_errorbarh(aes(xmin = OR_lower, xmax = OR_upper), height = 0.2) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  labs(
    title = "Bayesian Final Model",
    subtitle = "Odds ratios with 95% credible intervals for m = 10",
    x = "Odds Ratio",
    y = "Variable"
  ) +
  theme_minimal()

bayesian_forest_plot

ggsave(
  file.path(fig_path, "bayesian_forest_plot.png"),
  bayesian_forest_plot,
  width = 9,
  height = 5.5,
  dpi = 300
)

# Create predicted probability plot

pred_data <- map_dfr(
  fits_final_m10,
  function(fit) {
    data.frame(
      observed = fit$data$HeartDisease,
      predicted = fitted(fit)[, "Estimate"]
    )
  }
) %>%
  mutate(
    observed = factor(
      as.numeric(as.character(observed)),
      levels = c(0, 1),
      labels = c("No Disease", "Disease")
    )
  )

bayesian_predicted_plot <- ggplot(
  pred_data,
  aes(x = predicted, fill = observed)
) +
  geom_density(alpha = 0.4) +
  labs(
    title = "Bayesian Predicted Probabilities",
    subtitle = "Final Bayesian model for m = 10",
    x = "Predicted probability",
    y = "Density",
    fill = "Observed outcome"
  ) +
  theme_minimal()

bayesian_predicted_plot

ggsave(
  file.path(fig_path, "bayesian_predicted.png"),
  bayesian_predicted_plot,
  width = 9,
  height = 5.5,
  dpi = 300
)
```
