---
title: "06 - Cross Validation"
output:  html_document
---

# Overview

This script evaluates and compares the predictive performance of the final frequentist and Bayesian logistic regression models using cross-validation.

The script:
- loads a representative completed dataset from the multiple imputation procedure,
- performs stratified 10-fold cross-validation for the frequentist and Bayesian models,
- computes predictive performance measures (AUC, Accuracy, Sensitivity, and Specificity),
- generates ROC curves based on held-out predictions,
- compares frequentist 5-fold, 10-fold, and leave-one-out cross-validation,
- evaluates the computational runtime of the cross-validation procedures,
- exports summary tables and figures used in the thesis.

```{r}

# ==================================================
# Load packages, paths, and imputed dataset
# ==================================================

library(mice)
library(dplyr)
library(purrr)
library(tibble)
library(tidyr)
library(ggplot2)
library(caret)
library(pROC)
library(brms)
library(posterior)
library(knitr)
library(kableExtra)

options(brms.backend = "rstan")

# Paths

input_path <- "Imputed-Data"

cv_results_path <- file.path(
  "Results",
  "Cross Validation"
)

fig_path <- "figures"

dir.create(cv_results_path, recursive = TRUE, showWarnings = FALSE)
dir.create(fig_path, recursive = TRUE, showWarnings = FALSE)

# Load imputed dataset

imp_m5 <- readRDS(
  file.path(input_path, "imputed_data_m5.rds")
)

# Use one representative completed dataset from m = 5

data_cv <- complete(imp_m5, 1)

data_cv <- data_cv %>%
  mutate(
    HeartDisease = as.numeric(as.character(HeartDisease))
  )

```



```{r}
# ==================================================
# Define final models and Bayesian settings
# ==================================================

# Final frequentist model
final_formula_freq <- HeartDisease ~ Age + Sex + ChestPainType +
  FastingBS + ExerciseAngina + Oldpeak + ST_Slope

# Final Bayesian model
final_formula_bayes <- HeartDisease ~ Sex + ChestPainType +
  FastingBS + ExerciseAngina + Oldpeak + ST_Slope

# Priors 
priors <- c(
  prior(student_t(7, 0, 2.5), class = "b"),
  prior(student_t(7, 0, 10), class = "Intercept")
)

# HMC settings 
bayes_chains <- 4
bayes_cores <- min(4, parallel::detectCores())
bayes_iter <- 800
bayes_warmup <- 400
bayes_seed <- 123
```



```{r}
# ==================================================
# Create stratified 10-fold cross-validation folds
# ==================================================

set.seed(123)

folds <- createFolds(
  y = factor(data_cv$HeartDisease),
  k = 10,
  list = TRUE,
  returnTrain = FALSE
)

# Check class distribution in each fold
fold_distribution <- map_dfr(
  seq_along(folds),
  function(i) {
    
    test_data <- data_cv[folds[[i]], ]
    
    tibble(
      fold = i,
      n = nrow(test_data),
      no_disease = sum(test_data$HeartDisease == 0),
      disease = sum(test_data$HeartDisease == 1),
      disease_prevalence = mean(test_data$HeartDisease == 1)
    )
  }
)

fold_distribution

write.csv(
  fold_distribution,
  file.path(cv_results_path, "stratified_10_fold_distribution.csv"),
  row.names = FALSE
)
```


```{r}
# ==================================================
# Define predictive performance helper functions
# ==================================================

compute_cv_performance <- function(pred_data) {
  
  pred_data <- pred_data %>%
    mutate(
      predicted_class = ifelse(predicted_probability >= 0.5, 1, 0)
    )
  
  TP <- sum(pred_data$observed == 1 & pred_data$predicted_class == 1)
  TN <- sum(pred_data$observed == 0 & pred_data$predicted_class == 0)
  FP <- sum(pred_data$observed == 0 & pred_data$predicted_class == 1)
  FN <- sum(pred_data$observed == 1 & pred_data$predicted_class == 0)
  
  roc_object <- roc(
    response = pred_data$observed,
    predictor = pred_data$predicted_probability
  )
  
  tibble(
    AUC = as.numeric(auc(roc_object)),
    Accuracy = mean(pred_data$predicted_class == pred_data$observed),
    Sensitivity = TP / (TP + FN),
    Specificity = TN / (TN + FP),
    TP = TP,
    TN = TN,
    FP = FP,
    FN = FN
  )
}

# Helper function for Confidence Intervals

mean_ci <- function(x) {
  
  x <- x[!is.na(x)]
  
  m <- mean(x)
  se <- sd(x) / sqrt(length(x))
  
  c(
    mean = m,
    lower = m - 1.96 * se,
    upper = m + 1.96 * se
  )
}

format_mean_ci <- function(x, digits = 3) {
  
  ci <- mean_ci(x)
  
  paste0(
    round(ci["mean"], digits),
    " (",
    round(ci["lower"], digits),
    ", ",
    round(ci["upper"], digits),
    ")"
  )
}
```


```{r}
# ==================================================
# Run frequentist stratified 10-fold cross-validation
# ==================================================

# Frequentist stratified 10-fold cross-validation function
run_frequentist_cv <- function(data, folds, formula_final) {
  
  map_dfr(
    seq_along(folds),
    function(i) {
      
      test_index <- folds[[i]]
      
      train_data <- data[-test_index, ]
      test_data  <- data[test_index, ]
      
      fit <- glm(
        formula = formula_final,
        data = train_data,
        family = binomial
      )
      
      pred_prob <- predict(
        fit,
        newdata = test_data,
        type = "response"
      )
      
      tibble(
        fold = i,
        observed = test_data$HeartDisease,
        predicted_probability = pred_prob,
        method = "Frequentist"
      )
    }
  )
}

  
freq_cv_predictions <- run_frequentist_cv(
    data = data_cv,
    folds = folds,
    formula_final = final_formula_freq
  )
  
freq_cv_predictions

write.csv(
  freq_cv_predictions,
  file.path(cv_results_path, "frequentist_10_fold_cv_predictions.csv"),
  row.names = FALSE
)
```


```{r}
# ==================================================
# Summarise frequentist cross-validation performance
# ==================================================

freq_cv_fold_performance <- freq_cv_predictions %>%
  group_by(fold) %>%
  group_modify(~ compute_cv_performance(.x)) %>%
  ungroup() %>%
  mutate(method = "Frequentist")

freq_cv_summary <- freq_cv_fold_performance %>%
  summarise(
    method = "Frequentist",
    AUC = format_mean_ci(AUC),
    Accuracy = format_mean_ci(Accuracy),
    Sensitivity = format_mean_ci(Sensitivity),
    Specificity = format_mean_ci(Specificity)
  )

freq_cv_fold_performance
freq_cv_summary
```

```{r}
# ==================================================
# Run Bayesian stratified 10-fold cross-validation
# ==================================================

# Bayesian stratified 10-fold cross-validation function
run_bayesian_cv <- function(data, folds, formula_final) {
  
  map_dfr(
    seq_along(folds),
    function(i) {
      
      cat("\nRunning Bayesian CV fold", i, "of", length(folds), "\n")
      
      test_index <- folds[[i]]
      
      train_data <- data[-test_index, ]
      test_data  <- data[test_index, ]
      
      fit <- brm(
        formula = formula_final,
        data = train_data,
        family = bernoulli(link = "logit"),
        prior = priors,
        chains = bayes_chains,
        iter = bayes_iter,
        warmup = bayes_warmup,
        cores = bayes_cores,
        seed = bayes_seed,
        control = list(adapt_delta = 0.95),
        refresh = 0
      )
      
      pred_prob <- fitted(
        fit,
        newdata = test_data
      )[, "Estimate"]
      
      tibble(
        fold = i,
        observed = test_data$HeartDisease,
        predicted_probability = pred_prob,
        method = "Bayesian"
      )
    }
  )
}

bayes_10fold_time <- system.time({
  
  bayes_cv_predictions <- run_bayesian_cv(
    data = data_cv,
    folds = folds,
    formula_final = final_formula_bayes
  )
  
})

bayes_10fold_time
bayes_cv_predictions

write.csv(
  bayes_cv_predictions,
  file.path(cv_results_path, "bayesian_10_fold_cv_predictions.csv"),
  row.names = FALSE
)
```

```{r}
# ==================================================
# Summarise Bayesian cross-validation performance
# ==================================================

bayes_cv_fold_performance <- bayes_cv_predictions %>%
  group_by(fold) %>%
  group_modify(~ compute_cv_performance(.x)) %>%
  ungroup() %>%
  mutate(method = "Bayesian")

bayes_cv_summary <- bayes_cv_fold_performance %>%
  summarise(
    method = "Bayesian",
    AUC = format_mean_ci(AUC),
    Accuracy = format_mean_ci(Accuracy),
    Sensitivity = format_mean_ci(Sensitivity),
    Specificity = format_mean_ci(Specificity)
  )

bayes_cv_fold_performance
bayes_cv_summary
```

```{r}
# ==================================================
# Combine and save cross-validation results
# ==================================================

cv_predictions_all <- bind_rows(
  freq_cv_predictions,
  bayes_cv_predictions
)

cv_fold_performance_all <- bind_rows(
  freq_cv_fold_performance,
  bayes_cv_fold_performance
)

cv_summary_all <- bind_rows(
  freq_cv_summary,
  bayes_cv_summary
)

cv_predictions_all
cv_fold_performance_all
cv_summary_all

write.csv(
  cv_predictions_all,
  file.path(cv_results_path, "stratified_10_fold_cv_predictions_all.csv"),
  row.names = FALSE
)

write.csv(
  cv_fold_performance_all,
  file.path(cv_results_path, "stratified_10_fold_cv_fold_performance.csv"),
  row.names = FALSE
)

write.csv(
  cv_summary_all,
  file.path(cv_results_path, "stratified_10_fold_cv_summary_CI.csv"),
  row.names = FALSE
)
```



```{r}
# ==================================================
# Create ROC curves from held-out predictions
# ==================================================

roc_freq_cv <- roc(
  response = freq_cv_predictions$observed,
  predictor = freq_cv_predictions$predicted_probability
)

roc_bayes_cv <- roc(
  response = bayes_cv_predictions$observed,
  predictor = bayes_cv_predictions$predicted_probability
)

# Convert ROC objects to data frames

roc_freq_df <- tibble(
  FalsePositiveRate = 1 - roc_freq_cv$specificities,
  TruePositiveRate = roc_freq_cv$sensitivities
)

roc_bayes_df <- tibble(
  FalsePositiveRate = 1 - roc_bayes_cv$specificities,
  TruePositiveRate = roc_bayes_cv$sensitivities
)

# AUC values

auc_freq <- round(as.numeric(auc(roc_freq_cv)), 3)
auc_bayes <- round(as.numeric(auc(roc_bayes_cv)), 3)

# =========================================================
# Frequentist ROC curve
# =========================================================

freq_roc_plot <- ggplot(
  roc_freq_df,
  aes(
    x = FalsePositiveRate,
    y = TruePositiveRate
  )
) +
  
  geom_line(
    linewidth = 1.4,
    colour = "black"
  ) +
  
  geom_abline(
    intercept = 0,
    slope = 1,
    linetype = "dashed",
    colour = "grey60"
  ) +
  
  coord_cartesian(
    xlim = c(0, 1),
    ylim = c(0, 1)
  ) +
  
  labs(
    title = "Frequentist ROC Curve",
    subtitle = paste0(
      "Stratified 10-Fold Cross-Validation (AUC = ",
      auc_freq,
      ")"
    ),
    x = "1 - Specificity",
    y = "Sensitivity"
  ) +
  
  theme_minimal(base_size = 15) +
  
  theme(
    plot.title = element_text(
      face = "bold",
      size = 18
    ),
    
    plot.subtitle = element_text(
      size = 14
    ),
    
    axis.title = element_text(
      face = "bold",
      size = 15
    ),
    
    axis.text = element_text(
      size = 13
    ),
    
    panel.grid.minor = element_blank()
  )

freq_roc_plot

ggsave(
  filename = file.path(
    fig_path,
    "frequentist_cv_roc_curve.png"
  ),
  plot = freq_roc_plot,
  width = 8,
  height = 5.5,
  dpi = 300
)

# =========================================================
# Bayesian ROC curve
# =========================================================

bayes_roc_plot <- ggplot(
  roc_bayes_df,
  aes(
    x = FalsePositiveRate,
    y = TruePositiveRate
  )
) +
  
  geom_line(
    linewidth = 1.4,
    colour = "black"
  ) +
  
  geom_abline(
    intercept = 0,
    slope = 1,
    linetype = "dashed",
    colour = "grey60"
  ) +
  
  coord_cartesian(
    xlim = c(0, 1),
    ylim = c(0, 1)
  ) +
  
  labs(
    title = "Bayesian ROC Curve",
    subtitle = paste0(
      "Stratified 10-Fold Cross-Validation (AUC = ",
      auc_bayes,
      ")"
    ),
    x = "1 - Specificity",
    y = "Sensitivity"
  ) +
  
  theme_minimal(base_size = 15) +
  
  theme(
    plot.title = element_text(
      face = "bold",
      size = 18
    ),
    
    plot.subtitle = element_text(
      size = 14
    ),
    
    axis.title = element_text(
      face = "bold",
      size = 15
    ),
    
    axis.text = element_text(
      size = 13
    ),
    
    panel.grid.minor = element_blank()
  )

bayes_roc_plot

ggsave(
  filename = file.path(
    fig_path,
    "bayesian_cv_roc_curve.png"
  ),
  plot = bayes_roc_plot,
  width = 8,
  height = 5.5,
  dpi = 300
)
```


```{r}
# ==================================================
# Define function for frequentist CV method comparison
# ==================================================

run_frequentist_cv_comparison <- function(data, formula_final, k = NULL, method_name) {
  
  if (!is.null(k)) {
    
    set.seed(123)
    
    folds <- createFolds(
      y = factor(data$HeartDisease),
      k = k,
      list = TRUE,
      returnTrain = FALSE
    )
    
  } else {
    
    folds <- as.list(seq_len(nrow(data)))
  }
  
  predictions <- map_dfr(
    seq_along(folds),
    function(i) {
      
      test_index <- folds[[i]]
      
      train_data <- data[-test_index, ]
      test_data  <- data[test_index, ]
      
      fit <- glm(
        formula = formula_final,
        data = train_data,
        family = binomial
      )
      
      pred_prob <- predict(
        fit,
        newdata = test_data,
        type = "response"
      )
      
      tibble(
        fold = i,
        observed = test_data$HeartDisease,
        predicted_probability = pred_prob,
        method = method_name
      )
    }
  )
  
  predictions
}
```


```{r}
# ==================================================
# Run frequentist 5-fold, 10-fold, and LOO CV
# ==================================================

  
  freq_cv_5_predictions <- run_frequentist_cv_comparison(
    data = data_cv,
    formula_final = final_formula_freq,
    k = 5,
    method_name = "5-fold CV"
  )
  
  
  freq_cv_10_predictions <- run_frequentist_cv_comparison(
    data = data_cv,
    formula_final = final_formula_freq,
    k = 10,
    method_name = "10-fold CV"
  )
 


  
  freq_cv_loo_predictions <- run_frequentist_cv_comparison(
    data = data_cv,
    formula_final = final_formula_freq,
    k = NULL,
    method_name = "LOO CV"
  )
  

freq_cv_5_predictions
freq_cv_10_predictions
freq_cv_loo_predictions
```


```{r}
# ==================================================
# Compute fold-wise performance for 5-fold and 10-fold CV
# ==================================================

freq_cv_5_fold_performance <- freq_cv_5_predictions %>%
  group_by(fold) %>%
  group_modify(~ compute_cv_performance(.x)) %>%
  ungroup() %>%
  mutate(CV_Method = "5-fold CV")

freq_cv_10_fold_performance <- freq_cv_10_predictions %>%
  group_by(fold) %>%
  group_modify(~ compute_cv_performance(.x)) %>%
  ungroup() %>%
  mutate(CV_Method = "10-fold CV")

freq_cv_5_fold_performance
freq_cv_10_fold_performance
```

```{r}
# ==================================================
# Summarise 5-fold and 10-fold CV performance
# ==================================================

summarise_cv_mean_ci <- function(fold_performance, method_name) {
  
  fold_performance %>%
    summarise(
      CV_Method = method_name,
      AUC = format_mean_ci(AUC),
      Accuracy = format_mean_ci(Accuracy),
      Sensitivity = format_mean_ci(Sensitivity),
      Specificity = format_mean_ci(Specificity)
    )
}

freq_cv_5_summary <- summarise_cv_mean_ci(
  freq_cv_5_fold_performance,
  "5-fold CV"
)

freq_cv_10_summary <- summarise_cv_mean_ci(
  freq_cv_10_fold_performance,
  "10-fold CV"
)

freq_cv_5_summary
freq_cv_10_summary
```


```{r}
# ==================================================
# Compute overall LOO CV performance
# ==================================================

freq_cv_loo_performance <- compute_cv_performance(
  freq_cv_loo_predictions
)

freq_cv_loo_summary <- freq_cv_loo_performance %>%
  transmute(
    CV_Method = "LOO CV",
    AUC = as.character(round(AUC, 3)),
    Accuracy = as.character(round(Accuracy, 3)),
    Sensitivity = as.character(round(Sensitivity, 3)),
    Specificity = as.character(round(Specificity, 3))
  )

freq_cv_loo_performance
freq_cv_loo_summary
```


```{r}
# ==================================================
# Combine and save frequentist CV method comparison
# ==================================================

freq_cv_method_comparison <- bind_rows(
  freq_cv_5_summary,
  freq_cv_10_summary,
  freq_cv_loo_summary
)

freq_cv_method_comparison

write.csv(
  freq_cv_method_comparison,
  file.path(
    cv_results_path,
    "frequentist_5fold_10fold_loo_cv_comparison.csv"
  ),
  row.names = FALSE
)
```


```{r}
# ==================================================
# Save fold-wise frequentist CV performance
# ==================================================

freq_cv_fold_performance_comparison <- bind_rows(
  freq_cv_5_fold_performance,
  freq_cv_10_fold_performance
)

write.csv(
  freq_cv_fold_performance_comparison,
  file.path(
    cv_results_path,
    "frequentist_5fold_10fold_fold_performance.csv"
  ),
  row.names = FALSE
)
```



```{r}
# ==================================================
# Evaluate cross-validation runtime stability
# ==================================================

n_time_runs <- 5

time_freq_10fold_cv <- function(run_id) {
  
  set.seed(1000 + run_id)
  
  time_out <- system.time({
    run_frequentist_cv(
      data = data_cv,
      folds = folds,
      formula_final = final_formula_freq
    )
  })
  
  tibble(
    Procedure = "Frequentist stratified 10-fold CV",
    run = run_id,
    elapsed_time_seconds = as.numeric(time_out["elapsed"])
  )
}

time_bayes_10fold_cv <- function(run_id) {
  
  set.seed(1000 + run_id)
  
  time_out <- system.time({
    run_bayesian_cv(
      data = data_cv,
      folds = folds,
      formula_final = final_formula_bayes
    )
  })
  
  tibble(
    Procedure = "Bayesian stratified 10-fold CV",
    run = run_id,
    elapsed_time_seconds = as.numeric(time_out["elapsed"])
  )
}

time_freq_cv_comparison <- function(run_id, k_value, method_name) {
  
  set.seed(1000 + run_id)
  
  time_out <- system.time({
    run_frequentist_cv_comparison(
      data = data_cv,
      formula_final = final_formula_freq,
      k = k_value,
      method_name = method_name
    )
  })
  
  tibble(
    Procedure = method_name,
    run = run_id,
    elapsed_time_seconds = as.numeric(time_out["elapsed"])
  )
}

cv_runtime_stability_raw <- bind_rows(
  map_dfr(1:n_time_runs, time_freq_10fold_cv),
  map_dfr(1:n_time_runs, time_bayes_10fold_cv),
  map_dfr(1:n_time_runs, ~ time_freq_cv_comparison(.x, 5, "Frequentist 5-fold CV")),
  map_dfr(1:n_time_runs, ~ time_freq_cv_comparison(.x, 10, "Frequentist 10-fold CV")),
  map_dfr(1:n_time_runs, ~ time_freq_cv_comparison(.x, NULL, "Frequentist LOO CV"))
)

cv_runtime_stability_summary <- cv_runtime_stability_raw %>%
  group_by(Procedure) %>%
  summarise(
    Mean_Runtime = paste0(
      round(mean(elapsed_time_seconds, na.rm = TRUE), 3),
      " ± ",
      round(sd(elapsed_time_seconds, na.rm = TRUE), 3),
      " sec"
    ),
    .groups = "drop"
  )

cv_runtime_stability_summary

write.csv(
  cv_runtime_stability_raw,
  file.path(cv_results_path, "cross_validation_runtime_stability_raw.csv"),
  row.names = FALSE
)

write.csv(
  cv_runtime_stability_summary,
  file.path(cv_results_path, "cross_validation_runtime_stability_summary.csv"),
  row.names = FALSE
)
```


