---
title: "03 - Frequentist Logistic Regression"
output: html_document
---
# Overview

This script implements the frequentist logistic regression analysis used in the thesis.

The script:
- loads the imputed datasets,
- fits full logistic regression models,
- performs backward variable selection using pooled p-values,
- compares models across imputation settings,
- computes predicted probabilities and predictive performance,
- evaluates computational runtime,
- generates figures.

```{r}
# ==================================================
# Load packages, paths, and imputed datasets
# ==================================================

library(mice)
library(dplyr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(grid)
library(pROC)
library(caret)
library(purrr)

input_path <- "Imputed-Data"

freq_results_path <- file.path(
  "Results",
  "Frequentist Approach",
  "Pooling results"
)

fig_path <- "figures"

dir.create(freq_results_path, recursive = TRUE, showWarnings = FALSE)
dir.create(fig_path, recursive = TRUE, showWarnings = FALSE)

imp_m5 <- readRDS(file.path(input_path, "imputed_data_m5.rds"))
imp_m10 <- readRDS(file.path(input_path, "imputed_data_m10.rds"))
imp_m20 <- readRDS(file.path(input_path, "imputed_data_m20.rds"))

```



```{r}
# ==================================================
# Define predictors and significance-star function 
# ==================================================

predictors_full <- c(
  "Age", "Sex", "ChestPainType", "RestingBP", "Cholesterol",
  "FastingBS", "RestingECG", "MaxHR", "ExerciseAngina",
  "Oldpeak", "ST_Slope"
)

add_significance_stars <- function(results) {
  results %>%
    mutate(
      significance = case_when(
        p.value < 0.001 ~ "***",
        p.value < 0.01  ~ "**",
        p.value < 0.05  ~ "*",
        TRUE            ~ ""
      ),
      OR_with_stars = paste0(round(estimate, 2), significance)
    )
}
```


```{r}
# ==================================================
# Define functions for pooled frequentist models
# ==================================================

fit_full_pooled_model <- function(imp_object) {
  
  fit <- with(
    data = imp_object,
    expr = glm(
      HeartDisease ~ Age + Sex + ChestPainType + RestingBP + Cholesterol +
        FastingBS + RestingECG + MaxHR + ExerciseAngina + Oldpeak + ST_Slope,
      family = binomial
    )
  )
  
  pooled_fit <- pool(fit)
  
  summary(
    pooled_fit,
    conf.int = TRUE,
    exponentiate = TRUE
  ) %>%
    add_significance_stars()
}

# Fit and pool the logistic regression model across imputations
fit_pool_model <- function(imp_object, predictors) {
  
  form <- as.formula(
    paste("HeartDisease ~", paste(predictors, collapse = " + "))
  )
  
  completed_data <- complete(imp_object, action = "all")
  
  fit_list <- lapply(completed_data, function(dat) {
    glm(form, data = dat, family = binomial)
  })
  
  fit <- as.mira(fit_list)
  pooled <- pool(fit)
  
  results <- summary(
    pooled,
    conf.int = TRUE,
    exponentiate = TRUE
  ) %>%
    add_significance_stars()
  
  results_sorted <- results %>%
    filter(term != "(Intercept)") %>%
    arrange(desc(p.value))
  
  list(
    formula = form,
    fit = fit,
    pooled = pooled,
    results = results,
    results_sorted = results_sorted
  )
}
```



```{r}
# ==================================================
# Define backward elimination using pooled p-values
# ==================================================

get_base_predictor <- function(term) {
  case_when(
    grepl("^ChestPainType", term) ~ "ChestPainType",
    grepl("^RestingECG", term) ~ "RestingECG",
    grepl("^ST_Slope", term) ~ "ST_Slope",
    grepl("^Sex", term) ~ "Sex",
    grepl("^FastingBS", term) ~ "FastingBS",
    grepl("^ExerciseAngina", term) ~ "ExerciseAngina",
    TRUE ~ term
  )
}

# Perform backward variable selection using pooled p-values
backward_selection_pooled <- function(imp_object, predictors, p_cutoff = 0.05) {
  
  current_predictors <- predictors
  removed_variables <- c()
  selection_steps <- data.frame()
  step <- 0
  
  repeat {
    step <- step + 1
    
    model_out <- fit_pool_model(imp_object, current_predictors)
    res_sorted <- model_out$results_sorted
    
    worst_term <- res_sorted$term[1]
    worst_p <- res_sorted$p.value[1]
    remove_var <- get_base_predictor(worst_term)
    
    selection_steps <- rbind(
      selection_steps,
      data.frame(
        step = step,
        model_formula = paste(deparse(model_out$formula), collapse = ""),
        worst_term = worst_term,
        removed_candidate = remove_var,
        p_value = worst_p,
        predictors_before_removal = paste(current_predictors, collapse = ", ")
      )
    )
    
    cat("\nStep", step, "\n")
    cat("Formula:", deparse(model_out$formula), "\n")
    cat("Worst term:", worst_term, "\n")
    cat("P-value:", worst_p, "\n")
    
    if (worst_p <= p_cutoff) {
      cat("Stopping: all remaining terms have p <=", p_cutoff, "\n")
      break
    }
    
    cat("Removed variable:", remove_var, "\n")
    
    removed_variables <- c(removed_variables, remove_var)
    current_predictors <- setdiff(current_predictors, remove_var)
  }
  
  final_model <- fit_pool_model(imp_object, current_predictors)
  
  list(
    final_predictors = current_predictors,
    removed_variables = removed_variables,
    selection_steps = selection_steps,
    final_model = final_model
  )
}
```


```{r}
# ==================================================
# Fit full frequentist models 
# ==================================================

results_full_m5 <- fit_full_pooled_model(imp_m5)

results_full_m10 <- fit_full_pooled_model(imp_m10)

results_full_m20 <- fit_full_pooled_model(imp_m20)

```



```{r}
# ==================================================
# Run backward elimination
# ==================================================

bw_manual_m5 <- backward_selection_pooled(
    imp_object = imp_m5,
    predictors = predictors_full,
    p_cutoff = 0.05
  )
  

bw_manual_m10 <- backward_selection_pooled(
    imp_object = imp_m10,
    predictors = predictors_full,
    p_cutoff = 0.05
  )
  
bw_manual_m20 <- backward_selection_pooled(
    imp_object = imp_m20,
    predictors = predictors_full,
    p_cutoff = 0.05
  )

```


```{r}
# ==================================================
# Save full model and selected model results
# ==================================================

write.csv(
  results_full_m5,
  file.path(freq_results_path, "frequentist_full_results_m5.csv"),
  row.names = FALSE
)

write.csv(
  results_full_m10,
  file.path(freq_results_path, "frequentist_full_results_m10.csv"),
  row.names = FALSE
)

write.csv(
  results_full_m20,
  file.path(freq_results_path, "frequentist_full_results_m20.csv"),
  row.names = FALSE
)

write.csv(
  bw_manual_m5$selection_steps,
  file.path(freq_results_path, "backward_selection_steps_m5.csv"),
  row.names = FALSE
)

write.csv(
  bw_manual_m10$selection_steps,
  file.path(freq_results_path, "backward_selection_steps_m10.csv"),
  row.names = FALSE
)

write.csv(
  bw_manual_m20$selection_steps,
  file.path(freq_results_path, "backward_selection_steps_m20.csv"),
  row.names = FALSE
)

write.csv(
  bw_manual_m5$final_model$results,
  file.path(freq_results_path, "frequentist_selected_results_m5.csv"),
  row.names = FALSE
)

write.csv(
  bw_manual_m10$final_model$results,
  file.path(freq_results_path, "frequentist_selected_results_m10.csv"),
  row.names = FALSE
)

write.csv(
  bw_manual_m20$final_model$results,
  file.path(freq_results_path, "frequentist_selected_results_m20.csv"),
  row.names = FALSE
)
```


```{r}
# ==================================================
# Compare final selected models across imputations
# ==================================================

extract_final_summary <- function(bw_object, m_value) {
  
  final_results <- bw_object$final_model$results
  removed_vars <- setdiff(predictors_full, bw_object$final_predictors)
  
  final_results %>%
    mutate(
      m = m_value,
      selected_variables = paste(bw_object$final_predictors, collapse = ", "),
      number_selected_variables = length(bw_object$final_predictors),
      removed_variables = paste(removed_vars, collapse = ", "),
      number_removed_variables = length(removed_vars)
    )
}

comparison_final_results <- bind_rows(
  extract_final_summary(bw_manual_m5, 5),
  extract_final_summary(bw_manual_m10, 10),
  extract_final_summary(bw_manual_m20, 20)
) %>%
  select(
    m,
    term,
    estimate,
    OR_with_stars,
    std.error,
    statistic,
    df,
    p.value,
    significance,
    conf.low,
    conf.high,
    selected_variables,
    number_selected_variables,
    removed_variables,
    number_removed_variables
  )

write.csv(
  comparison_final_results,
  file.path(freq_results_path, "comparison_final_results_m5_m10_m20.csv"),
  row.names = FALSE
)
```


```{r}
# ======================================================
# Compare backward elimination steps across imputations
# ======================================================

extract_selection_summary <- function(bw_object, m_value) {
  
  bw_object$selection_steps %>%
    mutate(m = m_value) %>%
    select(
      m,
      step,
      worst_term,
      removed_candidate,
      p_value,
      predictors_before_removal
    )
}

comparison_selection_steps <- bind_rows(
  extract_selection_summary(bw_manual_m5, 5),
  extract_selection_summary(bw_manual_m10, 10),
  extract_selection_summary(bw_manual_m20, 20)
)


write.csv(
  comparison_selection_steps,
  file.path(freq_results_path, "comparison_selection_steps_m5_m10_m20.csv"),
  row.names = FALSE
)
```


```{r}
# ============================================================
# Compute predicted probabilities for final frequentist models
# ============================================================

get_frequentist_predicted_probabilities <- function(bw_object, imp_object, m_value) {
  
  final_formula <- bw_object$final_model$formula
  completed_data <- mice::complete(imp_object, action = "all")
  
  pred_data <- lapply(seq_along(completed_data), function(i) {
    
    dat <- completed_data[[i]]
    
    fit <- glm(
      formula = final_formula,
      data = dat,
      family = binomial
    )
    
    data.frame(
      imputation = i,
      m = m_value,
      observed = dat$HeartDisease,
      predicted_probability = predict(fit, type = "response")
    )
  })
  
  bind_rows(pred_data)
}


freq_pred_m5 <- get_frequentist_predicted_probabilities(
    bw_manual_m5,
    imp_m5,
    5
  )
  

freq_pred_m10 <- get_frequentist_predicted_probabilities(
    bw_manual_m10,
    imp_m10,
    10
  )
  

freq_pred_m20 <- get_frequentist_predicted_probabilities(
    bw_manual_m20,
    imp_m20,
    20
  )
  

freq_pred_all <- bind_rows(
  freq_pred_m5,
  freq_pred_m10,
  freq_pred_m20
)

```

```{r}
# ===============================================================
# Average frequentist predicted probabilities across imputations
# ===============================================================

average_frequentist_predictions <- function(pred_data, m_value) {
  
  pred_data %>%
    group_by(imputation) %>%
    mutate(observation = row_number()) %>%
    ungroup() %>%
    group_by(observation, observed) %>%
    summarise(
      predicted_probability = mean(predicted_probability),
      .groups = "drop"
    ) %>%
    mutate(m = m_value)
}

freq_pred_avg_m5 <- average_frequentist_predictions(freq_pred_m5, 5)
freq_pred_avg_m10 <- average_frequentist_predictions(freq_pred_m10, 10)
freq_pred_avg_m20 <- average_frequentist_predictions(freq_pred_m20, 20)

write.csv(
  freq_pred_avg_m5,
  file.path(freq_results_path, "frequentist_predicted_probabilities_avg_m5.csv"),
  row.names = FALSE
)

write.csv(
  freq_pred_avg_m10,
  file.path(freq_results_path, "frequentist_predicted_probabilities_avg_m10.csv"),
  row.names = FALSE
)

write.csv(
  freq_pred_avg_m20,
  file.path(freq_results_path, "frequentist_predicted_probabilities_avg_m20.csv"),
  row.names = FALSE
)
```



```{r}
# ============================================================
# Frequentist runtime stability: mean ± SD over repeated runs
# ============================================================

n_time_runs <- 5

time_frequentist_full_model <- function(run_id, imp_object, m_value) {
  
  time_out <- system.time({
    fit_full_pooled_model(imp_object)
  })
  
  tibble(
    procedure = "Full model",
    m = m_value,
    run = run_id,
    elapsed_time_seconds = as.numeric(time_out["elapsed"])
  )
}

time_frequentist_backward_selection <- function(run_id, imp_object, m_value) {
  
  time_out <- system.time({
    backward_selection_pooled(
      imp_object = imp_object,
      predictors = predictors_full,
      p_cutoff = 0.05
    )
  })
  
  tibble(
    procedure = "Backward selection",
    m = m_value,
    run = run_id,
    elapsed_time_seconds = as.numeric(time_out["elapsed"])
  )
}

time_frequentist_final_model <- function(run_id, bw_object, imp_object, m_value) {
  
  time_out <- system.time({
    fit_pool_model(
      imp_object = imp_object,
      predictors = bw_object$final_predictors
    )
  })
  
  tibble(
    procedure = "Final selected model",
    m = m_value,
    run = run_id,
    elapsed_time_seconds = as.numeric(time_out["elapsed"])
  )
}

time_frequentist_prediction <- function(run_id, bw_object, imp_object, m_value) {
  
  time_out <- system.time({
    get_frequentist_predicted_probabilities(
      bw_object = bw_object,
      imp_object = imp_object,
      m_value = m_value
    )
  })
  
  tibble(
    procedure = "Predicted probabilities",
    m = m_value,
    run = run_id,
    elapsed_time_seconds = as.numeric(time_out["elapsed"])
  )
}

freq_runtime_stability_raw <- bind_rows(
  
  map_dfr(1:n_time_runs, ~ time_frequentist_full_model(.x, imp_m5, 5)),
  
  map_dfr(1:n_time_runs, ~ time_frequentist_backward_selection(.x, imp_m5, 5)),
  
  map_dfr(1:n_time_runs, ~ time_frequentist_final_model(.x, bw_manual_m5, imp_m5, 5)),
  
  map_dfr(1:n_time_runs, ~ time_frequentist_prediction(.x, bw_manual_m5, imp_m5, 5))
)

freq_runtime_stability_summary <- freq_runtime_stability_raw %>%
  group_by(procedure, m) %>%
  summarise(
    Mean_Runtime = paste0(
      round(mean(elapsed_time_seconds, na.rm = TRUE), 3),
      " ± ",
      round(sd(elapsed_time_seconds, na.rm = TRUE), 3),
      " sec"
    ),
    .groups = "drop"
  )

write.csv(
  freq_runtime_stability_raw,
  file.path(freq_results_path, "frequentist_runtime_stability_raw.csv"),
  row.names = FALSE
)

write.csv(
  freq_runtime_stability_summary,
  file.path(freq_results_path, "frequentist_runtime_stability_summary.csv"),
  row.names = FALSE
)
```


```{r}
# ============================================================
# Save frequentist predicted probabilities
# ============================================================

write.csv(
  freq_pred_m5,
  file.path(freq_results_path, "frequentist_predicted_probabilities_m5.csv"),
  row.names = FALSE
)

write.csv(
  freq_pred_m10,
  file.path(freq_results_path, "frequentist_predicted_probabilities_m10.csv"),
  row.names = FALSE
)

write.csv(
  freq_pred_m20,
  file.path(freq_results_path, "frequentist_predicted_probabilities_m20.csv"),
  row.names = FALSE
)

write.csv(
  freq_pred_all,
  file.path(freq_results_path, "frequentist_predicted_probabilities_all.csv"),
  row.names = FALSE
)
```


```{r}
# ============================================================
# Plot frequentist predicted probabilities for m = 10
# ============================================================

freq_pred_plot_m10 <- freq_pred_m10 %>%
  mutate(
    observed = factor(
      observed,
      levels = c(0, 1),
      labels = c("No Disease", "Disease")
    )
  ) %>%
  ggplot(aes(x = predicted_probability, fill = observed)) +
  geom_density(alpha = 0.4) +
  labs(
    title = "Frequentist Predicted Probabilities",
    subtitle = "Final frequentist logistic regression model for m = 10",
    x = "Predicted probability",
    y = "Density",
    fill = "Observed outcome"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )

freq_pred_plot_m10

ggsave(
  filename = file.path(fig_path, "frequentist_predicted_probabilities_m10.png"),
  plot = freq_pred_plot_m10,
  width = 9,
  height = 5.5,
  dpi = 300
)
```

```{r}
# =================================================================
# Frequentist predictive performance with 95% confidence intervals
# =================================================================

compute_frequentist_performance_by_imputation <- function(pred_data, m_value) {
  
  pred_data <- pred_data %>%
    mutate(
      observed = as.numeric(as.character(observed)),
      predicted_class = ifelse(predicted_probability >= 0.5, 1, 0)
    )
  
  pred_data %>%
    group_by(imputation) %>%
    summarise(
      AUC = as.numeric(
        auc(
          roc(
            response = observed,
            predictor = predicted_probability,
            quiet = TRUE
          )
        )
      ),
      Accuracy = mean(predicted_class == observed),
      
      TP = sum(predicted_class == 1 & observed == 1),
      TN = sum(predicted_class == 0 & observed == 0),
      FP = sum(predicted_class == 1 & observed == 0),
      FN = sum(predicted_class == 0 & observed == 1),
      
      Sensitivity = ifelse((TP + FN) == 0, NA, TP / (TP + FN)),
      Specificity = ifelse((TN + FP) == 0, NA, TN / (TN + FP)),
      
      .groups = "drop"
    ) %>%
    mutate(m = m_value)
}


mean_ci <- function(x) {
  
  x <- x[!is.na(x)]
  
  m <- mean(x)
  se <- sd(x) / sqrt(length(x))
  
  c(
    mean = m,
    lower = m - 1.96 * se,
    upper = m + 1.96 * se
  )
}


summarise_frequentist_performance <- function(performance_by_imp, m_value) {
  
  performance_by_imp %>%
    summarise(
      m = m_value,
      n_imputations = n(),
      
      AUC_mean = mean_ci(AUC)["mean"],
      AUC_lower = mean_ci(AUC)["lower"],
      AUC_upper = mean_ci(AUC)["upper"],
      
      Accuracy_mean = mean_ci(Accuracy)["mean"],
      Accuracy_lower = mean_ci(Accuracy)["lower"],
      Accuracy_upper = mean_ci(Accuracy)["upper"],
      
      Sensitivity_mean = mean_ci(Sensitivity)["mean"],
      Sensitivity_lower = mean_ci(Sensitivity)["lower"],
      Sensitivity_upper = mean_ci(Sensitivity)["upper"],
      
      Specificity_mean = mean_ci(Specificity)["mean"],
      Specificity_lower = mean_ci(Specificity)["lower"],
      Specificity_upper = mean_ci(Specificity)["upper"]
    )
}

# Compute performance for each imputed dataset

freq_performance_by_imp_m5 <- compute_frequentist_performance_by_imputation(
  pred_data = freq_pred_m5,
  m_value = 5
)

freq_performance_by_imp_m10 <- compute_frequentist_performance_by_imputation(
  pred_data = freq_pred_m10,
  m_value = 10
)

freq_performance_by_imp_m20 <- compute_frequentist_performance_by_imputation(
  pred_data = freq_pred_m20,
  m_value = 20
)


# Summarise mean performance with 95% confidence intervals

freq_performance_summary_m5 <- summarise_frequentist_performance(
  freq_performance_by_imp_m5,
  m_value = 5
)

freq_performance_summary_m10 <- summarise_frequentist_performance(
  freq_performance_by_imp_m10,
  m_value = 10
)

freq_performance_summary_m20 <- summarise_frequentist_performance(
  freq_performance_by_imp_m20,
  m_value = 20
)


freq_performance_results <- bind_rows(
  freq_performance_summary_m5,
  freq_performance_summary_m10,
  freq_performance_summary_m20
)

```


```{r} 
# ===============================
# Plot ROC curve for m = 10
# ===============================
freq_roc_m10 <- roc(
  response = freq_pred_avg_m10$observed,
  predictor = freq_pred_avg_m10$predicted_probability,
  quiet = TRUE
)

roc_plot_freq_m10 <- ggroc(
  freq_roc_m10,
  legacy.axes = TRUE,
  linewidth = 1.4
) +
  geom_abline(
    intercept = 0,
    slope = 1,
    linetype = "dashed",
    colour = "grey60"
  ) +
  labs(
    title = "Frequentist ROC Curve",
    subtitle = paste0(
      "Final frequentist model for m = 10 (AUC = ",
      round(freq_performance_summary_m10$AUC_mean, 3),
      ")"
    ),
    x = "1 - Specificity",
    y = "Sensitivity"
  ) +
  coord_cartesian(
    xlim = c(0, 1),
    ylim = c(0, 1)
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(
      face = "bold",
      size = 18
    ),
    
    plot.subtitle = element_text(
      size = 14
    ),
    
    axis.title = element_text(
      face = "bold",
      size = 15
    ),
    
    axis.text = element_text(
      size = 13
    ),
    
    panel.grid.minor = element_blank()
  )

roc_plot_freq_m10

ggsave(
  filename = file.path(
    fig_path,
    "frequentist_roc_curve_m10.png"
  ),
  plot = roc_plot_freq_m10,
  width = 8,
  height = 5.5,
  dpi = 300
)
```


```{r}
# ============================================================
# Save frequentist predictive performance results
# ============================================================

write.csv(
  freq_performance_by_imp_m5,
  file.path(freq_results_path, "frequentist_predictive_performance_by_imputation_m5.csv"),
  row.names = FALSE
)

write.csv(
  freq_performance_by_imp_m10,
  file.path(freq_results_path, "frequentist_predictive_performance_by_imputation_m10.csv"),
  row.names = FALSE
)

write.csv(
  freq_performance_by_imp_m20,
  file.path(freq_results_path, "frequentist_predictive_performance_by_imputation_m20.csv"),
  row.names = FALSE
)

write.csv(
  freq_performance_results,
  file.path(freq_results_path, "frequentist_predictive_performance_summary_CI.csv"),
  row.names = FALSE
)
```



```{r}
# =========================
# Create frequentist plots
# =========================

selection_plot_data <- comparison_selection_steps %>%
  mutate(
    m = factor(m, levels = c(5, 10, 20)),
    removed_candidate = factor(
      removed_candidate,
      levels = unique(removed_candidate)
    ),
    p_label = ifelse(p_value < 0.001, "<0.001", round(p_value, 3))
  )

backward_selection_plot <- ggplot(
  selection_plot_data,
  aes(x = step, y = p_value, group = m)
) +
  geom_hline(yintercept = 0.05, linetype = "dashed") +
  geom_line(linewidth = 1) +
  geom_point(size = 3.5) +
  geom_text(
  aes(label = removed_candidate),
  hjust = 0,
  nudge_x = 0.12,
  size = 3.5,
  vjust = 0
) +
  scale_x_continuous(
  limits = c(1, 6),
  breaks = 1:6,
  expand = expansion(mult = c(0.02, 0.02))
) +
  facet_wrap(~ m, labeller = label_both) +
  labs(
    title = "Backward Selection Based on Pooled p-values",
    x = "Backward Selection Step",
    y = "Largest Pooled p-value"
  ) +
  theme_minimal(base_size = 16) +
 theme(
  plot.title = element_text(
    face = "bold",
    size = 18,
    hjust = 0.5
  ),
  axis.title = element_text(
    size = 16,
    face = "bold"
  ),
  axis.text = element_text(
    size = 14
  ),
  strip.text = element_text(
    size = 15,
    face = "bold"
  ),
  panel.grid.minor = element_blank(),

  # Space between facets
  panel.spacing = unit(1.5, "cm")
)

backward_selection_plot

# Save plot

ggsave(
  filename = file.path(fig_path, "frequentist_backward_selection_plot.png"),
  plot = backward_selection_plot,
  width = 10,
  height = 7,
  dpi = 300
)
```




