---
title: "02 - Multiple Imputation"
output: html_document
---

# Overview

This script performs the missing-data preprocessing and multiple imputation used in the thesis.

The script:
- recodes implausible zero values in `RestingBP` and `Cholesterol` as missing,
- applies Multiple Imputation by Chained Equations (MICE),
- generates imputed datasets for \(m = 5, 10, 20\),
- saves the imputed objects and completed datasets,
- produces diagnostic plots for the imputation procedure.

```{r}
# ==================================================
# Load packages, paths, and data
# ==================================================

library(tidyverse)
library(knitr)
library(mice)

data_path <- "../heart disease dataset.csv"
output_path <- "Imputed-Data"

fig_path <- file.path(output_path, "figures")

if (!dir.exists(output_path)) {
  dir.create(output_path, recursive = TRUE)
}

if (!dir.exists(fig_path)) {
  dir.create(fig_path, recursive = TRUE)
}

data <- read_csv(data_path)
```


```{r}
# ==================================================
# Prepare variables and recode implausible zeros
# ==================================================

data <- data %>%
  mutate(
    Sex = factor(Sex),
    ChestPainType = factor(ChestPainType),
    FastingBS = factor(FastingBS),
    RestingECG = factor(RestingECG),
    ExerciseAngina = factor(ExerciseAngina),
    ST_Slope = factor(ST_Slope),
    HeartDisease = factor(HeartDisease),
    RestingBP = ifelse(RestingBP == 0, NA, RestingBP),
    Cholesterol = ifelse(Cholesterol == 0, NA, Cholesterol)
  )

colSums(is.na(data))

```

```{r, fig.width=15}
# ==================================================
# Inspect missing data and define MICE settings
# ==================================================

par(mar = c(5, 4, 4, 2), cex.axis = 1.5)
md.pattern(data)
par(cex.axis = 1.1)

init <- mice(data, maxit = 0, printFlag = FALSE)

meth <- init$method
pred <- init$predictorMatrix

meth["RestingBP"] <- "pmm"
meth["Cholesterol"] <- "pmm"

meth[c(
  "Age", "Sex", "ChestPainType", "FastingBS",
  "RestingECG", "MaxHR", "ExerciseAngina",
  "Oldpeak", "ST_Slope", "HeartDisease"
)] <- ""
```


```{r}
# ==================================================
# Run MICE for m = 5, 10, and 20
# ==================================================

run_mice <- function(m_value, seed_value = 123) {
  mice(
    data,
    m = m_value,
    method = meth,
    predictorMatrix = pred,
    maxit = 10,
    seed = seed_value,
    printFlag = FALSE
  )
}

imp_m5  <- run_mice(m_value = 5, seed_value = 123)
imp_m10 <- run_mice(m_value = 10, seed_value = 123)
imp_m20 <- run_mice(m_value = 20, seed_value = 123)

```


```{r}
# ==================================================
# Save MICE objects and completed datasets
# ==================================================

saveRDS(imp_m5,  file.path(output_path, "imputed_data_m5.rds"))
saveRDS(imp_m10, file.path(output_path, "imputed_data_m10.rds"))
saveRDS(imp_m20, file.path(output_path, "imputed_data_m20.rds"))

save_completed_datasets <- function(imp_object, m_value) {
  imputed_list <- complete(imp_object, action = "all")
  
  for (i in seq_along(imputed_list)) {
    write.csv(
      imputed_list[[i]],
      file = file.path(output_path, paste0("imputed_dataset_m", m_value, "_", i, ".csv")),
      row.names = FALSE
    )
  }
}

save_completed_datasets(imp_m5, 5)
save_completed_datasets(imp_m10, 10)
save_completed_datasets(imp_m20, 20)
```


```{r}
# ==================================================
# Examine imputation diagnostics
# ==================================================

plot(imp_m5)
densityplot(imp_m5, ~ Cholesterol, lwd = 2)
stripplot(imp_m5, Cholesterol ~ .imp, pch = 20)
stripplot(imp_m5, RestingBP ~ .imp, pch = 20)

plot(imp_m10)
densityplot(imp_m10, ~ Cholesterol, lwd = 2)
stripplot(imp_m10, Cholesterol ~ .imp, pch = 20)
stripplot(imp_m10, RestingBP ~ .imp, pch = 20)

plot(imp_m20)
densityplot(imp_m20, ~ Cholesterol, lwd = 2)
stripplot(imp_m20, Cholesterol ~ .imp, pch = 20)
stripplot(imp_m20, RestingBP ~ .imp, pch = 20)
```


```{r}
# ==================================================
# Save diagnostic plots
# ==================================================

save_imputation_plots <- function(imp_object, m_value) {
  
  png(file.path(fig_path, paste0("traceplot_m", m_value, ".png")),
      width = 1200, height = 800, res = 150)
  plot(imp_object)
  dev.off()
  
  png(file.path(fig_path, paste0("density_cholesterol_m", m_value, ".png")),
      width = 1200, height = 800, res = 150)
  print(densityplot(imp_object, ~ Cholesterol, lwd = 2))
  dev.off()
  
  png(file.path(fig_path, paste0("stripplot_cholesterol_m", m_value, ".png")),
      width = 1200, height = 800, res = 150)
  print(stripplot(imp_object, Cholesterol ~ .imp, pch = 20))
  dev.off()
  
  png(file.path(fig_path, paste0("stripplot_restingbp_m", m_value, ".png")),
      width = 1200, height = 800, res = 150)
  print(stripplot(imp_object, RestingBP ~ .imp, pch = 20))
  dev.off()
}

save_imputation_plots(imp_m5, 5)
save_imputation_plots(imp_m10, 10)
save_imputation_plots(imp_m20, 20)
```

