---
title: "07 - Subsampling"
output: html_document
---

# Overview

This script evaluates the stability of the frequentist and Bayesian modelling procedures under reduced sample sizes.

The script:
- loads one completed dataset from the \(m = 5\) imputation setting,
- creates repeated subsamples at 50%, 25%, 12.5%, and 6.25% of the original sample size,
- evaluates variable-selection stability for frequentist and Bayesian methods,
- compares predictive performance under subsampling,
- summarises odds ratios and uncertainty intervals,
- evaluates runtime stability,
- exports figures and result tables.

```{r}
# ==================================================
# Load packages, paths, and imputed dataset
# ==================================================

library(mice)
library(posterior)
library(dplyr)
library(purrr)
library(tibble)
library(tidyr)
library(ggplot2)
library(brms)
library(nimble)
library(pROC)

options(brms.backend = "rstan")

# Paths

input_path <- "Imputed-Data"

subsampling_results_path <- file.path(
  "Results",
  "Subsampling Analysis"
)

fig_path <- "figures"

dir.create(subsampling_results_path, recursive = TRUE, showWarnings = FALSE)
dir.create(fig_path, recursive = TRUE, showWarnings = FALSE)

imp_m5 <- readRDS(
  file.path(input_path, "imputed_data_m5.rds")
)

data_full <- complete(imp_m5, 1)

```


```{r}
# ==================================================
# Define predictors and helper functions
# ==================================================

predictors_full <- c(
  "Age", "Sex", "ChestPainType", "RestingBP", "Cholesterol",
  "FastingBS", "RestingECG", "MaxHR", "ExerciseAngina",
  "Oldpeak", "ST_Slope"
)

all_original_vars <- predictors_full

get_base_predictor <- function(term) {
  
  term <- gsub("^b_", "", term)
  
  case_when(
    grepl("^ChestPainType", term) ~ "ChestPainType",
    grepl("^RestingECG", term) ~ "RestingECG",
    grepl("^ST_Slope", term) ~ "ST_Slope",
    grepl("^Sex", term) ~ "Sex",
    grepl("^FastingBS", term) ~ "FastingBS",
    grepl("^ExerciseAngina", term) ~ "ExerciseAngina",
    TRUE ~ term
  )
}

clean_variable_names <- function(terms) {
  terms %>%
    gsub("^Sex.*", "Sex", .) %>%
    gsub("^ChestPainType.*", "ChestPainType", .) %>%
    gsub("^FastingBS.*", "FastingBS", .) %>%
    gsub("^RestingECG.*", "RestingECG", .) %>%
    gsub("^ExerciseAngina.*", "ExerciseAngina", .) %>%
    gsub("^ST_Slope.*", "ST_Slope", .) %>%
    unique()
}

```


```{r}
# ==================================================
# Create subsamples for variable-selection stability
# ==================================================

set.seed(123)

B_freq <- 25
B_bayes <- 25

subsample_props <- c(0.50, 0.25, 0.125, 0.0625)

create_subsamples <- function(data, prop, B) {
  
  n_sub <- floor(nrow(data) * prop)
  
  map(
    1:B,
    ~ data[sample(seq_len(nrow(data)), size = n_sub, replace = FALSE), ]
  )
}

set.seed(123)

freq_subsamples <- list(
  p50  = create_subsamples(data_full, 0.50, B_freq),
  p25  = create_subsamples(data_full, 0.25, B_freq),
  p125 = create_subsamples(data_full, 0.125, B_freq),
  p625 = create_subsamples(data_full, 0.0625, B_freq)
)

set.seed(123)

bayes_subsamples <- list(
  p50  = create_subsamples(data_full, 0.50, B_bayes),
  p25  = create_subsamples(data_full, 0.25, B_bayes),
  p125 = create_subsamples(data_full, 0.125, B_bayes),
  p625 = create_subsamples(data_full, 0.0625, B_bayes)
)

```


```{r}
# ==================================================
# Define predictive performance function
# ==================================================

compute_classification_metrics <- function(data, probabilities, cutoff = 0.5) {
  
  observed <- as.numeric(as.character(data$HeartDisease))
  predicted_class <- ifelse(probabilities >= cutoff, 1, 0)
  
  roc_object <- pROC::roc(
    response = observed,
    predictor = probabilities,
    quiet = TRUE
  )
  
  TP <- sum(predicted_class == 1 & observed == 1)
  TN <- sum(predicted_class == 0 & observed == 0)
  FP <- sum(predicted_class == 1 & observed == 0)
  FN <- sum(predicted_class == 0 & observed == 1)
  
  tibble(
    AUC = as.numeric(pROC::auc(roc_object)),
    Accuracy = mean(predicted_class == observed),
    Sensitivity = ifelse((TP + FN) == 0, NA, TP / (TP + FN)),
    Specificity = ifelse((TN + FP) == 0, NA, TN / (TN + FP))
  )
}
```


```{r}
# ==================================================
# Define frequentist backward elimination function
# ==================================================

fit_glm_model <- function(data, predictors) {
  
  if (length(predictors) == 0) {
    form <- HeartDisease ~ 1
  } else {
    form <- as.formula(
      paste("HeartDisease ~", paste(predictors, collapse = " + "))
    )
  }
  
  fit <- suppressWarnings(
    glm(
      form,
      data = data,
      family = binomial,
      control = glm.control(maxit = 100)
    )
  )
  
  fit
}

backward_selection_single <- function(data, predictors, p_cutoff = 0.05) {
  
  current_predictors <- predictors
  removed_variables <- c()
  
  repeat {
    
    fit <- fit_glm_model(data, current_predictors)
    coef_table <- summary(fit)$coefficients
    
    results <- data.frame(
      term = rownames(coef_table),
      p_value = coef_table[, "Pr(>|z|)"],
      row.names = NULL
    ) %>%
      filter(term != "(Intercept)") %>%
      mutate(base_predictor = get_base_predictor(term))
    
    predictor_pvalues <- results %>%
      group_by(base_predictor) %>%
      summarise(max_p_value = max(p_value, na.rm = TRUE), .groups = "drop") %>%
      arrange(desc(max_p_value))
    
    worst_var <- predictor_pvalues$base_predictor[1]
    worst_p <- predictor_pvalues$max_p_value[1]
    
    if (is.na(worst_p) || worst_p <= p_cutoff) {
      break
    }
    
    removed_variables <- c(removed_variables, worst_var)
    current_predictors <- setdiff(current_predictors, worst_var)
    
    if (length(current_predictors) == 0) {
      break
    }
  }
  
  final_fit <- fit_glm_model(data, current_predictors)
  
  list(
    final_predictors = current_predictors,
    removed_variables = removed_variables,
    final_fit = final_fit
  )
}
```



```{r}
# ==================================================
# Run frequentist subsampling selection
# ==================================================

run_frequentist_subsampling <- function(subsamples, prop_label) {
  
  map_dfr(
    seq_along(subsamples),
    function(b) {
      
      cat("\nFrequentist subsampling:", prop_label,
          "| repetition", b, "of", length(subsamples), "\n")
      
      dat <- subsamples[[b]]
      
      selected_model <- backward_selection_single(
        data = dat,
        predictors = predictors_full,
        p_cutoff = 0.05
      )
      
      selected_vars <- selected_model$final_predictors
      
      coef_df <- data.frame(
        term = names(coef(selected_model$final_fit)),
        estimate = as.numeric(coef(selected_model$final_fit)),
        row.names = NULL
      ) %>%
        filter(term != "(Intercept)") %>%
        mutate(
          base_predictor = get_base_predictor(term),
          repetition = b,
          subsample = prop_label
        )
      
      tibble(
        repetition = b,
        subsample = prop_label,
        selected_variables = paste(selected_vars, collapse = ", "),
        number_selected = length(selected_vars),
        coefs = list(coef_df),
        Converged = selected_model$final_fit$converged
      )
    }
  )
}

freq_subsampling_summary <- bind_rows(
  run_frequentist_subsampling(freq_subsamples$p50, "50%"),
  run_frequentist_subsampling(freq_subsamples$p25, "25%"),
  run_frequentist_subsampling(freq_subsamples$p125, "12.5%"),
  run_frequentist_subsampling(freq_subsamples$p625, "6.25%")
)

freq_subsampling_summary
```



```{r}
# ==================================================
# Compute frequentist inclusion frequencies
# ==================================================

extract_inclusion_frequency <- function(summary_data, method_name) {
  
  map_dfr(
    seq_len(nrow(summary_data)),
    function(i) {
      
      selected <- unlist(strsplit(summary_data$selected_variables[i], ", "))
      
      tibble(
        variable = all_original_vars,
        selected = ifelse(all_original_vars %in% selected, 1, 0),
        repetition = summary_data$repetition[i],
        subsample = summary_data$subsample[i]
      )
    }
  ) %>%
    group_by(subsample, variable) %>%
    summarise(
      inclusion_frequency = mean(selected),
      .groups = "drop"
    ) %>%
    mutate(method = method_name)
}

freq_inclusion_frequencies <- extract_inclusion_frequency(
  freq_subsampling_summary,
  "Frequentist"
)

freq_inclusion_frequencies
```



```{r}
# ==================================================
# Define Bayesian settings and spike-and-slab selection
# ==================================================

priors <- c(
  prior(student_t(7, 0, 2.5), class = "b"),
  prior(student_t(7, 0, 10), class = "Intercept")
)

bayes_chains <- 4
bayes_cores <- min(4, parallel::detectCores())
bayes_iter <- 800
bayes_warmup <- 400


# Bayesian spike-and-slab variable selection 

full_formula_selection <- HeartDisease ~ Age + Sex + ChestPainType + RestingBP + Cholesterol +
  FastingBS + RestingECG + MaxHR + ExerciseAngina + Oldpeak + ST_Slope

run_spike_slab_selection <- function(dat, formula_full, niter = 4000, nburnin = 1000) {
  
  X <- model.matrix(formula_full, data = dat)[, -1]
  y <- as.numeric(as.character(dat$HeartDisease))
  
  N <- nrow(X)
  P <- ncol(X)
  
  code <- nimbleCode({
    for (i in 1:N) {
      logit(p[i]) <- beta0 + inprod(beta_included[1:P], X[i, 1:P])
      y[i] ~ dbern(p[i])
    }
    
    beta0 ~ dnorm(0, sd = 5)
    
    for (j in 1:P) {
      gamma[j] ~ dbern(0.5)
      beta[j] ~ dnorm(0, sd = 2.5)
      beta_included[j] <- gamma[j] * beta[j]
    }
  })
  
  model <- nimbleModel(
    code = code,
    constants = list(N = N, P = P, X = X),
    data = list(y = y),
    inits = list(beta0 = 0, beta = rep(0, P), gamma = rep(1, P))
  )
  
  conf <- configureMCMC(
    model,
    monitors = c("gamma", "beta_included")
  )
  
  mcmc <- buildMCMC(conf)
  
  cmodel <- compileNimble(model)
  cmcmc <- compileNimble(mcmc, project = model)
  
  samples <- runMCMC(
    cmcmc,
    niter = niter,
    nburnin = nburnin,
    samplesAsCodaMCMC = FALSE
  )
  
  gamma_samples <- samples[, grep("^gamma", colnames(samples))]
  beta_samples  <- samples[, grep("^beta_included", colnames(samples))]
  
  tibble(
    term = colnames(X),
    pip = colMeans(gamma_samples),
    posterior_mean_beta = colMeans(beta_samples),
    posterior_mean_OR = exp(colMeans(beta_samples))
  )
}
```



```{r}
# ==================================================
# Run Bayesian subsampling selection
# ==================================================


run_bayesian_subsampling_selection <- function(subsamples, prop_label) {
  
  map_dfr(
    seq_along(subsamples),
    function(b) {
      
      cat("\nBayesian subsampling:", prop_label,
          "| repetition", b, "of", length(subsamples), "\n")
      
      dat <- subsamples[[b]]
      
      bayes_out <- run_spike_slab_selection(
        dat = dat,
        formula_full = full_formula_selection,
        niter = 4000,
        nburnin = 1000
      )
      
      selected_raw <- bayes_out %>%
        filter(pip > 0.5) %>%
        pull(term)
      
      selected_vars <- clean_variable_names(selected_raw)
      
      tibble(
        repetition = b,
        subsample = prop_label,
        selected_variables = paste(selected_vars, collapse = ", "),
        number_selected = length(selected_vars)
      )
    }
  )
}

bayes_subsampling_summary <- bind_rows(
  run_bayesian_subsampling_selection(bayes_subsamples$p50, "50%"),
  run_bayesian_subsampling_selection(bayes_subsamples$p25, "25%"),
  run_bayesian_subsampling_selection(bayes_subsamples$p125, "12.5%"),
  run_bayesian_subsampling_selection(bayes_subsamples$p625, "6.25%")
)

bayes_subsampling_summary


```

```{r}
# ==================================================
# Compute Bayesian inclusion frequencies
# ==================================================

bayes_inclusion_frequencies <- extract_inclusion_frequency(
  bayes_subsampling_summary,
  "Bayesian"
)

bayes_inclusion_frequencies


```



```{r}
# ==================================================
# Create subsamples for predictive performance analysis
# ==================================================

B_perf <- 5

set.seed(999)

performance_subsamples <- list(
  p50  = create_subsamples(data_full, 0.50, B_perf),
  p25  = create_subsamples(data_full, 0.25, B_perf),
  p125 = create_subsamples(data_full, 0.125, B_perf),
  p625 = create_subsamples(data_full, 0.0625, B_perf)
)
```


```{r}
# ==================================================
# Define predictive performance functions under subsampling
# ==================================================

run_frequentist_predictive_performance <- function(subsamples, prop_label) {
  
  map_dfr(
    seq_along(subsamples),
    function(b) {
      
      dat <- subsamples[[b]]
      
      selected_model <- backward_selection_single(
        data = dat,
        predictors = predictors_full,
        p_cutoff = 0.05
      )
      
      selected_vars <- selected_model$final_predictors
      final_fit <- selected_model$final_fit
      
      probabilities <- predict(
        final_fit,
        newdata = dat,
        type = "response"
      )
      
      perf <- compute_classification_metrics(
        data = dat,
        probabilities = probabilities
      )
      
      coef_table <- summary(final_fit)$coefficients
      
coef_results <- data.frame(
  term = rownames(coef_table),
  estimate = coef_table[, "Estimate"],
  row.names = NULL
) %>%
  filter(term != "(Intercept)") %>%
  mutate(
    term = gsub("^b_", "", term),
    OR = exp(estimate),
    method = "Frequentist",
    subsample = prop_label,
    repetition = b
  )
      
      perf %>%
        mutate(
          method = "Frequentist",
          subsample = prop_label,
          repetition = b,
          selected_variables = paste(selected_vars, collapse = ", "),
          number_selected = length(selected_vars),
          converged = final_fit$converged,
          coefficient_results = list(coef_results)
        )
    }
  )
}


fit_bayesian_selected_model <- function(dat, selected_vars, seed_value) {
  
  if (length(selected_vars) == 0) {
    form <- HeartDisease ~ 1
  } else {
    form <- as.formula(
      paste("HeartDisease ~", paste(selected_vars, collapse = " + "))
    )
  }
  
  brm(
    formula = form,
    data = dat,
    family = bernoulli(link = "logit"),
    prior = priors,
    chains = bayes_chains,
    iter = bayes_iter,
    warmup = bayes_warmup,
    cores = bayes_cores,
    seed = seed_value,
    refresh = 0,
    control = list(adapt_delta = 0.95)
  )
}

# Bayesian predictive performance function

run_bayesian_predictive_performance <- function(subsamples, prop_label) {
  
  map_dfr(
    seq_along(subsamples),
    function(b) {
      
      dat <- subsamples[[b]]
      
      bayes_out <- run_spike_slab_selection(
        dat = dat,
        formula_full = full_formula_selection,
        niter = 4000,
        nburnin = 1000
      )
      
      selected_raw <- bayes_out %>%
        filter(pip > 0.5) %>%
        pull(term)
      
      selected_vars <- clean_variable_names(selected_raw)
      
      fit <- fit_bayesian_selected_model(
        dat = dat,
        selected_vars = selected_vars,
        seed_value = 4000 + b
      )
      
      probabilities <- fitted(fit, scale = "response")[, "Estimate"]
      
      perf <- compute_classification_metrics(
        data = dat,
        probabilities = probabilities
      )
      
posterior_or_draws <- as_draws_df(fit) %>%
  as.data.frame() %>%
  select(starts_with("b_")) %>%
  pivot_longer(
    cols = everything(),
    names_to = "term",
    values_to = "estimate"
  ) %>%
  filter(term != "b_Intercept") %>%
  mutate(
    term = gsub("^b_", "", term),
    OR = exp(estimate),
    method = "Bayesian",
    subsample = prop_label,
    repetition = b
  )
      
      perf %>%
        mutate(
          method = "Bayesian",
          subsample = prop_label,
          repetition = b,
          selected_variables = paste(selected_vars, collapse = ", "),
          number_selected = length(selected_vars),
          coefficient_results = list(posterior_or_draws)
        )
    }
  )
}

```


```{r}
# ==================================================
# Run predictive performance analysis
# ==================================================

predictive_performance_raw <- bind_rows(
  
  run_frequentist_predictive_performance(performance_subsamples$p50, "50%"),
  run_frequentist_predictive_performance(performance_subsamples$p25, "25%"),
  run_frequentist_predictive_performance(performance_subsamples$p125, "12.5%"),
  run_frequentist_predictive_performance(performance_subsamples$p625, "6.25%"),
  
  run_bayesian_predictive_performance(performance_subsamples$p50, "50%"),
  run_bayesian_predictive_performance(performance_subsamples$p25, "25%"),
  run_bayesian_predictive_performance(performance_subsamples$p125, "12.5%"),
  run_bayesian_predictive_performance(performance_subsamples$p625, "6.25%")
)

predictive_performance_raw


# Extract coefficient / posterior OR results from the 5 fitted subsample models

subsample_or_raw <- predictive_performance_raw %>%
  select(coefficient_results) %>%
  unnest(coefficient_results)
```


```{r}
# ==================================================
# Summarise odds ratios and uncertainty intervals
# ==================================================

freq_subsample_or_summary <- subsample_or_raw %>%
  filter(method == "Frequentist") %>%
  mutate(log_OR = log(OR)) %>%
  group_by(subsample, term) %>%
  summarise(
    n_subsamples = n_distinct(repetition),

    mean_log_OR = mean(log_OR, na.rm = TRUE),
    sd_log_OR = sd(log_OR, na.rm = TRUE),
    se_log_OR = sd_log_OR / sqrt(n_subsamples),

    log_lower = mean_log_OR - 1.96 * se_log_OR,
    log_upper = mean_log_OR + 1.96 * se_log_OR,

    mean_OR = exp(mean_log_OR),
    OR_lower = exp(log_lower),
    OR_upper = exp(log_upper),

    .groups = "drop"
  ) %>%
  select(
    subsample,
    term,
    n_subsamples,
    mean_OR,
    OR_lower,
    OR_upper
  )

bayes_subsample_or_summary <- subsample_or_raw %>%
  filter(method == "Bayesian") %>%
  group_by(subsample, term) %>%
  summarise(
    n_subsamples = n_distinct(repetition),
    mean_OR = mean(OR, na.rm = TRUE),
    OR_lower = quantile(OR, 0.025, na.rm = TRUE),
    OR_upper = quantile(OR, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

subsample_or_summary <- bind_rows(
  freq_subsample_or_summary %>%
    mutate(method = "Frequentist", interval_type = "95% CI"),
  
  bayes_subsample_or_summary %>%
    mutate(method = "Bayesian", interval_type = "95% CrI")
) %>%
  arrange(method, subsample, term)

subsample_or_summary
```



```{r}
# ==================================================
# Summarise predictive performance metrics
# ==================================================

mean_ci <- function(x) {
  
  x <- x[!is.na(x)]
  
  m <- mean(x)
  se <- sd(x) / sqrt(length(x))
  
  c(
    mean = m,
    lower = m - 1.96 * se,
    upper = m + 1.96 * se
  )
}


predictive_performance_summary <- predictive_performance_raw %>%
  group_by(method, subsample) %>%
  summarise(
    n_subsamples = n(),
    
    AUC_mean = mean_ci(AUC)["mean"],
    AUC_lower = mean_ci(AUC)["lower"],
    AUC_upper = mean_ci(AUC)["upper"],
    
    Accuracy_mean = mean_ci(Accuracy)["mean"],
    Accuracy_lower = mean_ci(Accuracy)["lower"],
    Accuracy_upper = mean_ci(Accuracy)["upper"],
    
    Sensitivity_mean = mean_ci(Sensitivity)["mean"],
    Sensitivity_lower = mean_ci(Sensitivity)["lower"],
    Sensitivity_upper = mean_ci(Sensitivity)["upper"],
    
    Specificity_mean = mean_ci(Specificity)["mean"],
    Specificity_lower = mean_ci(Specificity)["lower"],
    Specificity_upper = mean_ci(Specificity)["upper"],
    
    .groups = "drop"
  )

predictive_performance_summary

```


```{r}

# ==================================================
# Combine frequentist and Bayesian inclusion frequencies
# ==================================================

all_inclusion_frequencies <- bind_rows(
  freq_inclusion_frequencies,
  bayes_inclusion_frequencies
) %>%
  arrange(method, subsample, desc(inclusion_frequency))

all_inclusion_frequencies


```



```{r}

# ==================================================
# Plot inclusion frequencies
# ==================================================

inclusion_frequency_plot <- all_inclusion_frequencies %>%
  mutate(
    variable = factor(variable, levels = rev(all_original_vars)),
    subsample = factor(subsample, levels = c("50%", "25%", "12.5%", "6.25%"))
  ) %>%
  ggplot(aes(x = variable, y = inclusion_frequency, fill = method)) +
  geom_col(position = "dodge") +
  coord_flip() +
  facet_wrap(~ subsample, ncol = 2) +
  labs(
    x = "Variable",
    y = "Inclusion Frequency",
    fill = "Method"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 10),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 18, face = "bold"),
    panel.grid.minor = element_blank()
  )

inclusion_frequency_plot

ggsave(
  filename = file.path(fig_path, "subsampling_inclusion_frequencies.png"),
  plot = inclusion_frequency_plot,
  width = 12,
  height = 8,
  dpi = 400
)
```

```{r}
# ==================================================
# Derive consensus models from inclusion frequencies
# ==================================================

stability_threshold <- 0.50

subsampling_consensus_models <- 
  all_inclusion_frequencies %>%
  filter(inclusion_frequency >= stability_threshold) %>%
  group_by(method, subsample) %>%
  summarise(
    consensus_variables = paste(variable, collapse = ", "),
    .groups = "drop"
  )

subsampling_consensus_models


```



```{r}
# ==================================================
# Summarise model size stability
# ==================================================

model_size_stability <- bind_rows(
  freq_subsampling_summary %>%
    select(repetition, subsample, number_selected) %>%
    mutate(method = "Frequentist"),
  
  bayes_subsampling_summary %>%
    select(repetition, subsample, number_selected) %>%
    mutate(method = "Bayesian")
)

model_size_summary <- model_size_stability %>%
  group_by(method, subsample) %>%
  summarise(
    mean_number_selected = mean(number_selected),
    sd_number_selected = sd(number_selected),
    min_number_selected = min(number_selected),
    max_number_selected = max(number_selected),
    .groups = "drop"
  )

model_size_summary


```


```{r}
# ==================================================
# Evaluate subsampling runtime stability
# ==================================================

B_time <- 10
n_time_runs <- 5

runtime_props <- list(
  "50%"   = 0.50,
  "25%"   = 0.25,
  "12.5%" = 0.125,
  "6.25%" = 0.0625
)

run_frequentist_runtime_once <- function(run_id, prop_label, prop_value) {
  
  set.seed(500 + run_id)
  
  subsamples_time <- create_subsamples(
    data_full,
    prop_value,
    B_time
  )
  
  start_time <- Sys.time()
  
  invisible(
    run_frequentist_subsampling(
      subsamples_time,
      prop_label
    )
  )
  
  end_time <- Sys.time()
  
  total_time <- as.numeric(
    difftime(end_time, start_time, units = "secs")
  )
  
  tibble(
    method = "Frequentist",
    subsample = prop_label,
    run = run_id,
    total_runtime_seconds = total_time,
    mean_runtime_per_subsample = total_time / B_time
  )
}

run_bayesian_runtime_once <- function(run_id, prop_label, prop_value) {
  
  set.seed(500 + run_id)
  
  subsamples_time <- create_subsamples(
    data_full,
    prop_value,
    B_time
  )
  
  start_time <- Sys.time()
  
  invisible(
    run_bayesian_subsampling_selection(
      subsamples_time,
      prop_label
    )
  )
  
  end_time <- Sys.time()
  
  total_time <- as.numeric(
    difftime(end_time, start_time, units = "secs")
  )
  
  tibble(
    method = "Bayesian",
    subsample = prop_label,
    run = run_id,
    total_runtime_seconds = total_time,
    mean_runtime_per_subsample = total_time / B_time
  )
}

runtime_stability_raw <- bind_rows(
  
  map_dfr(
    names(runtime_props),
    function(prop_label) {
      
      map_dfr(
        1:n_time_runs,
        ~ run_frequentist_runtime_once(
          run_id = .x,
          prop_label = prop_label,
          prop_value = runtime_props[[prop_label]]
        )
      )
    }
  ),
  
  map_dfr(
    names(runtime_props),
    function(prop_label) {
      
      map_dfr(
        1:n_time_runs,
        ~ run_bayesian_runtime_once(
          run_id = .x,
          prop_label = prop_label,
          prop_value = runtime_props[[prop_label]]
        )
      )
    }
  )
)

runtime_stability_summary <- runtime_stability_raw %>%
  group_by(method, subsample) %>%
  summarise(
    
    Total_Runtime = paste0(
      round(mean(total_runtime_seconds, na.rm = TRUE), 2),
      " ± ",
      round(sd(total_runtime_seconds, na.rm = TRUE), 2),
      " sec"
    ),
    
    Mean_Runtime_Per_Subsample = paste0(
      round(mean(mean_runtime_per_subsample, na.rm = TRUE), 2),
      " ± ",
      round(sd(mean_runtime_per_subsample, na.rm = TRUE), 2),
      " sec"
    ),
    
    .groups = "drop"
  )

runtime_stability_summary
```


```{r}
# ==================================================
# Save subsampling results
# ==================================================

write.csv(
  subsample_or_raw,
  file.path(subsampling_results_path, "individual_subsample_OR_raw.csv"),
  row.names = FALSE
)

write.csv(
  subsample_or_summary,
  file.path(subsampling_results_path, "individual_subsample_OR_summary_intervals.csv"),
  row.names = FALSE
)

write.csv(
  predictive_performance_raw,
  file.path(subsampling_results_path, "individual_subsample_predictive_performance_raw.csv"),
  row.names = FALSE
)

write.csv(
  predictive_performance_summary,
  file.path(subsampling_results_path, "individual_subsample_predictive_performance_summary_CI.csv"),
  row.names = FALSE
)

write.csv(
  runtime_stability_summary,
  file.path(subsampling_results_path, "subsampling_runtime_stability_summary.csv"),
  row.names = FALSE
)

write.csv(
  runtime_stability_raw,
  file.path(subsampling_results_path, "subsampling_runtime_stability_raw.csv"),
  row.names = FALSE)


write.csv(all_inclusion_frequencies,
  file.path(subsampling_results_path, "subsampling_inclusion_frequencies.csv"),
  row.names = FALSE
)

write.csv(model_size_summary,
  file.path(subsampling_results_path, "subsampling_model_size_summary.csv"),
  row.names = FALSE
)




```




